### 2015-02-04 ACMTransCL/HK_TOCL.tex
Initial commit

\documentclass[acmtocl]{acmsmall}

\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{arrows}
\usepackage{amssymb}
\usepackage{amsfonts,amsmath,latexsym,stmaryrd}
\usepackage{listings}
\usepackage{cite}
\input{macro}
\usepackage{alltt}
\usepackage[all]{xy}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage{booktabs}
\usetikzlibrary{shapes,snakes}
% 
% \usepackage{theorem}
% 
% \theorembodyfont{\rmfamily}
% \newtheorem{Thm}{Theorem}
% \newtheorem{Defn}[Thm]{Definition}
% \newtheorem{Ex}[Thm]{Example}
% \newtheorem{Lem}[Thm]{Lemma}
% \newtheorem{Prop}[Thm]{Proposition}
\newtheorem{PS}[theorem]{Proof Strategy}
\newtheorem{PEM}[theorem]{Proof Exploration Method}
% \newtheorem{GR}{General Remark}
% \newtheorem{Alg}{Algorithm}

\begin{document}

\markboth{J. Heras and E. Komendantskaya}{Proof-Pattern Search in Coq/SSReflect}

\title{Proof-Pattern Search in Coq/SSReflect}

%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{J\'onathan Heras
\affil{Department of Mathematics and Computer Science, University of La Rioja, Spain}
Ekaterina Komendantskaya
\affil{School of Computing, University of Dundee, UK}
}
%

\begin{abstract}
ML4PG is an extension of the Proof General interface, allowing
the user to invoke machine-learning algorithms and find proof similarities
in Coq/SSReflect libraries. In this paper, we present four new improvements to
%techniques that enhance the quality of the results obtained by 
ML4PG.
First, a new method of \lq\lq{}recurrent clustering" is introduced to collect statistical features from Coq terms.
Now, the user can receive suggestions about similar definitions,
types and lemma statements, in addition to proof strategies. Second, Coq
proofs are split into patches to capture proof strategies that could arise
at different stages of a proof. Third, we present a method to automatically generate Coq proofs based on ML4PG's output. Finally,
we introduce several visualisation tools that help to interpret the families of similar theorems/terms generated by ML4PG.
\end{abstract}

\category{F.4.1}{Mathematical Logic and Formal Languages}{Mathematical Logic -- Mechanical theorem proving}
\category{D.2.4}{Software Engineering}{Software/Program Verification -- Formal methods.}

\terms{}

\keywords{Coq/SSReflect, Proof Patterns, Recurrent Clustering, Pattern Recognition, Feature Extraction.}

\acmformat{J\'onathan Heras and Ekaterina Komendantskaya, 2014.
Proof-Pattern Search in Coq/SSReflect.}

\begin{bottomstuff}
The work was supported by EPSRC grants EP/J014222/1 and EP/K031864/1.

Author's addresses: J\'onathan Heras and Ekaterina Komendantskaya,
School of Computing, University of Dundee, UK
\end{bottomstuff}

\maketitle

\input{introduction}

\input{lemmaclustering}

\input{rec}

\input{patches}

\input{evaluation}

\input{visualisation}

\input{relatedwork}

\input{conclusions}

%\section{Acknowledgements}

\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{biblio}

\pagebreak

\appendix

\input{gallina}

\input{coqtactics}

\input{automaton}

\end{document}