### 2015-02-04 ACMTransCL/conceptualisation.tex

\section{Visualisation tools in ML4PG}\label{sec:visualisation}

In the previous section, we have seen how ML4PG can be used to find families of proofs following 
a common proof pattern. However, the output provided by clustering algorithms is just a set
of similar patches with no hints of why these proof-patches are deemed similar. 
% In addition, the user needs
% to replay the proof to infer the pattern. %Therefore, the conceptualisation of proof-patterns can be
% %a challenge. 
In this section, we present our approach to facilitate the understanding of proof patterns.

The first problem that we address is the discovery of the key features that were taken
into account during the cluster formation. This is a well-known problem in machine-learning 
known as \emph{feature selection}~\cite{Weka}. From a given set of features, feature selection algorithms 
generate different subsets of features and create a ranking of feature subsets.
ML4PG uses the \emph{correlation-based feature subset selection} algorithm implemented in Weka~\cite{Weka}
-- the machine-learning toolbox employed by ML4PG to extract the relevant features.

\begin{example}\label{ex:correlation-features}
In the case study presented at the end of the previous section, the relevant features were:

$-$ The tactics applied in the first and second steps of the proof-patches (\lstinline?case? and \lstinline?rewrite? respectively).

$-$ The type of the argument of the tactic applied in the first step of the proof-patches (\lstinline?nat?).

$-$ The second top symbol of the goal in the second step of the proof-patches (the $\sum$ symbol).

$-$ The first and second auxiliary lemmas applied in the the third step of the proof-patches. The first auxiliary lemma (\lstinline?sumrB?) is used to expand the summations and is 
common to all the proofs. There are two different \lq\lq{}second\rq\rq{} auxiliary lemmas that occur in the proofs of Lemma~\ref{lem:fundamental} and~\ref{lem:nilpotent2} (\lstinline?big_nat_recr? and
\lstinline?big_nat_recl? -- both extract elements of a summation); however, the recurrent feature extraction process have assigned similar values to them.
The ``automaton'' of Example~\ref{ex:correlation-features} is depicted in Figure~\ref{fig:automata} and Appendix~\ref{sec:automaton}.
\end{example}

ML4PG uses this information and produces an automaton-shape representation for discovered % to help in the visualisation of 
proof-patterns and the correlated features that determined the patterns. 
Our graphical representation is simpler than other works where automata are used to represent models that are inferred from proof traces~\cite{GWR14}.  
% This feature selection mechanism helps to explain why the proof-patches are similar since it indicates the 
% features that have been relevant for the construction of the cluster. To show the selected features and their relation with 
% the proof-patches, we introduce an automaton-shape representation. In~\cite{GWR14}, a tool to infer models 
% from proof traces was presented. Those models help to understand the behaviour exhibited by a collection
% of proofs and is represented by automata.

Generally, given a cluster of proof patches ${\cal C}$, we have an automaton $A$ with $5$ consecutive states. The $i$th state of $A$ is labelled
with the list of $i$th goals in the proof-patches contained in ${\cal C}$. The transitions between the $i$th and $i+1$th states are given by 
the $i$th tactic of each proof-patch of ${\cal C}$. If two or more tactics belong to the same group (see Figure~\ref{tab:tactics}),
they are merged in a unique transition; otherwise, the tactics will be shown as different transitions. In addition, each state is annotated with 
features 
whose correlation determined the cluster.
%for clustering that are related to that 
%state.

% $-$ there are several transitions from the $j$th to the $j+1$th state labelled by the $j$th tactic of each proof-patch of $P$ -- if either the $j$th tactic of two or more
% proof-patches is the same or belong to the same group (see Table~\ref{tab:tactics}), the transitions are merged in a unique transition labelled with the set of those tactics,
% 
% $-$ if a proof ends after the $j$th tactic of a proof-patch, the $j+1$ state is represented as final.

%\begin{example}

\begin{figure}[t]
\centering 
\includegraphics[scale=.19]{itp6.png}
\caption{\scriptsize{\emph{Fragment of the automaton 
 corresponding to the cluster of four lemma fragments described in the case study of Section~\ref{sec:recurrent}} (the whole automaton can be seen in Appendix~\ref{sec:automaton}). 
It shows correspondence between certain proof steps of lemmas % the following lemmas were discovered:
\texttt{telescope} (for the lemma of Example~\ref{example0}), \texttt{nilpotent\_inverse} (for Lemma~\ref{lem:nilpotent2}), and \texttt{persistent\_homology} 
(for Lemma~\ref{lem:fundamental}). Square boxes denote feature correlation where it exists. 
To be compact, we  hide full lemma statements, tactic and auxiliary lemma names, symbols, etc., but they can be shown by ML4PG. 
%For example, the full statements of the following cluster lemmas will be shown:
%The names of the lemmas of that cluster are: 
}}\label{fig:automata}
\end{figure}

%\end{example}