### 2015-02-04 ACMTransCL/evaluation.tex

\section{Evaluation}\label{sec:evaluation}

In the previous sections, we have shown how ML4PG has been designed to provide \emph{interesting} and \emph{non-trivial} hints on user's demand, 
and to be flexible enough to do so at any stage of the proof, and relative to any chosen proof library. 
However, it is difficult to measure how ML4PG improves the interactive proof building experience, since
the usability of a hint varies from user to user. In this section, we present a quantitative method to evaluate
how useful ML4PG can be during the proof development.

Machine-learning techniques have been previously used for the automatically generation of proofs in 
ITPs, see~\cite{Duncan02,DixonF03,GNR14}. ML4PG was not initially designed with this aim;
but we can use the information obtained from clustering to automatically generate proof attempts for a given theorem.

\begin{PEM}\label{pem}
Given the statement of a theorem $T$ and a set of lemmas $L=\{L_i\}_i$, we can use ML4PG to find a proof for $T$ as follows:
\begin{enumerate}
 \item Using term-clustering and $T\cup L$ as dataset, obtain the cluster $C$ that contains the theorem $T$ (i.e. $C=T\cup \{L_j\}_j$ where $\{L_j\}_j$ is the set of lemmas similar to $T$).
 \item Obtain the sequence of tactics $\{T_1^j,\ldots,T_{n_i}^j\}_{j}$ used to prove each lemma $L_j$ in $C$.
 \item For each $j$, try to prove $T$ using $T_1^j,\ldots,T_{n_j}^j$.
 \item If no sequence of tactics prove $T$, then for each tactic use ML4PG to infer the argument for each tactic $T_k^j$:
    \begin{itemize}
     \item If the argument of $T_k^j$ is an internal hypothesis from the context of a proof, try all the internal hypothesis from the context of the current proof. 
     \item If the argument of $T_k^j$ is an external lemma $E$, use term-clustering and $L$ as dataset to compute the lemmas in the same cluster as $E$ and try all those lemmas.
     \item[***] This can be naturally extended to tactics with several arguments, just trying all the possible combinations. 
  \end{itemize}    
\end{enumerate}
\end{PEM}

\begin{example}\label{ex:maxnACA}
Let $T$ be the lemma \lstinline?maxnACA? that states the inner commutativity of the maximum of two natural numbers in the SSReflect library:

\begin{lstlisting}
Lemma maxnACA : interchange maxn maxn. 
\end{lstlisting}

\noindent and $L$ be the \lstinline?ssrnat? library of SSReflect. Using Proof Exploration Method~\ref{pem}, we can construct a 
proof of \lstinline?maxnACA? as follows.

\begin{enumerate}
 \item $T$ belongs to the cluster containing the lemmas $\{$\lstinline?addnACA?, \lstinline?minnACA?, \lstinline?mulnACA?$\}$ --- these
 3 lemmas state the inner commutativity of addition, multiplication and the minimum of two naturals respectively. 
 \item For this example, we will only consider the proof of the lemma \lstinline?addnACA? that is proven using the sequence of tactics \lstinline?by move=> m n p q; rewrite -!addnA (addnCA n).?
 \item The proof of \lstinline?maxnACA? fails using the sequence of tactics from \lstinline?addnACA?.
 \item The proof of \lstinline?addnACA? uses the lemmas \lstinline?addnA? and \lstinline?addnCA?. If we cluster these lemmas with the
 rest of the lemmas of the ssrnat library we find the cluster $\{$\lstinline?minnA?, \lstinline?mulnA?, \lstinline?maxnA?$\}$ for 
 \lstinline?addnA?, and the cluster $\{$\lstinline?minnAC?, \lstinline?mulnAC?, \lstinline?maxnAC?$\}$ for 
 \lstinline?addnAC?. 
 \item Using the lemmas \lstinline?maxnA? and \lstinline?maxnAC?, we construct the sequence of tactics \lstinline?by move=> m n p q; rewrite -!maxnA (maxnCA n).? that proves the lemma \lstinline?maxnACA?.
\end{enumerate}

\end{example}

Using 5 Coq theories of varied sizes, we perform an empirical evaluation of our proof exploration method. Our test data consists of the Basic infrastructure of SSReflect library~\cite{SSReflect}, the formalisation about Java-like bytecode presented in~\cite{HK14}, the formalisation of persistent homology~\cite{HCMS12}, the paths library of the HoTT development~\cite{hottbook}, and two formalisations of Nash Equilibrium~\cite{Ves06,nash}. Using Proof exploration method~\ref{pem}, we try to prove every lemma of these libraries with a fully-honest approach (following the terminology from~\cite{KaliszykU14}): clustering is only performed against the lemmas that have previously
proven in the given library. In addition, we study the impact of changing the granularity value.

The results of our experiments can be seen in Table~\ref{tab:reproven}. The success rate of the proof exploration method 
depends on how similar are the proofs of theorems in a given library. This explains the high success rate in the \lstinline?Paths?
library, where most of the lemmas are proven using almost the same sequence of tactics, and the low rate in the Persistent Homology 
library, where just a few lemmas are similarly proven. The granularity value does not have a big impact in the performance of 
our experiments, and almost the same amount of lemmas are proven with the different values. In some cases, like in the Nash equilibrium
library, a small granularity value generates bigger clusters that increase the exploration space allowing to prove more lemmas. 
However, reducing the granularity value can also have a negative impact; for instance, in the JVM library the number of clusters
is reduced and this means a reduction of the proven theorems.

The evaluation method presented in this section shows an estimation of how ML4PG can help during the proof development. 
This evaluation is a bit artificial since it heavily relies on having a well-developed background theory. For instance, 
in Example~\ref{ex:maxnACA}, the proof exploration method is able to automatically generate a proof for \lstinline?maxnACA? since
the lemmas \lstinline?maxnA? and \lstinline?maxnAC? were already in the library, but this is not a common scenario when the user
is creating his library. However, ML4PG can be useful for the user even if those two lemmas are not available: ML4PG can suggest
that \lstinline?maxnACA? is similar to  \lstinline?addnACA?, and from the proof of the latter theorem, the user can infer that he needs
to state the lemmas \lstinline?maxnA? and \lstinline?maxnAC? that were missing in the library. The automatic generation of lemmas based
on clustering information was studied for the ACL2 system in~\cite{lpar13} --- the extrapolation of the technique presented in~\cite{lpar13} to Coq is not trivial since ACL2 works with first-order logic and Coq works with higher-order logic.

% 
% \begin{itemize}
%  \item Basic infrastructure of SSReflect library~\cite{SSReflect}: 1389 theorems formalised using SSReflect.
%  \item Formalisation about Java-like bytecode~\cite{HK14}: 49 theorems formalised using SSReflect.
%  \item Formalisation of Persistent Homology~\cite{HCMS12}: 306 theorems formalised using SSReflect.
%  \item Library Paths of HoTT development~\cite{hottbook}: 80 theorems formalised using plain Coq.
%  \item Formalisation of Nash Equilibrium~\cite{Ves06,nash}: 145 theorems formalised using plain Coq.
% \end{itemize}

\begin{table}
\tbl{\scriptsize{\emph{Percentage of automatically re-proven theorems.}}\label{tab:reproven}}{
\centering
\begin{tabular}{cccccc}
 \hline
 Library & Language & Granularity 1 & Granularity 3 & Granularity 5 & Theorems \\
 \hline

 SSReflect library  & SSReflect & $36\%$ & $35\%$ & $28\%$ & 1389\\

  JVM & SSReflect  & $56\%$ & $58\%$ & $65\%$ & 49\\

 Persistent Homology & SSReflect  & $0\%$ & $10\%$ & $12\%$ & 306\\

 Paths (HoTT) & Coq & $92\%$ & $91\%$ & $94\%$ &80\\

 Nash Equilibrium & Coq & $40\%$ & $37\%$ & $36\%$ & 145\\
   \hline
\end{tabular}}

\end{table}