### 2015-08-06 AI4FM/article.tex
Improvements

\documentclass{eceasst}
% This is an empty ECEASST article that can be used as a template
% by authors.
% Just uncomment the appropriate frontmatter commands and provide
% the parameters.

% Required packages
% =================
% Your \usepackage commands go here.

% Volume frontmatter
% ==================
\input{frontmatter}

% Article frontmatter
% ===================
\title{Scaling Automated Theory Exploration} % Title of the article
%\short{} % Short title of the article (optional)
\author{Chris Warburton\sponsor{}} % Authors and references to addresses
\institute{ % Institutes with labels
\email{cmwarburton@dundee.ac.uk}\\
School of Computing\\
University of Dundee}
\abstract{
We investigate the \textbf{theory exploration} (TE)
paradigm for computer-assisted Mathematics and identify limitations and
improvements for current approaches. Unlike the theorem-proving paradigm,
which requires user-provided conjectures, TE performs an open-ended
search for theorems satisfying given criteria. We see promise in TE for
identifying new abstractions and connections in libraries of software
and proofs, but realising this potential requires more scalable
algorithms than presently used.} % Abstract of the article
%\keywords{} % Keywords for the article

\begin{document}
\maketitle

% Main part of your article
% =========================
\section{Introduction}

The \emph{theory exploration} (TE) paradigm provides software support
for traditional Mathematical workflows; namely, deriving ``interesting''
consequences from formal definitions \cite{RISC1482}. Early
implementations like \textsc{Theorema} \cite{buchberger2000theory} emphasised
interactivity, in a similar way to computer algebra systems like
\textsc{Mathematica} (in which \textsc{Theorema} is implemented) or
interactive theorem provers. Subsequent systems have investigated
\emph{automated} theory exploration, for tasks such as lemma discovery
\cite{Hipster}.

By removing user interaction, automated TE systems require an algorithm
for deciding whether to inform the user of a particular theorem or not; we refer
to this search criterion as the theorem being ``interesting''.

In existing systems, this criterion is intimately connected to the
search algorithm; this coupling improves efficiency, even for brute-force
search. As an example, \textsc{IsaCoSy} \cite{johansson2009isacosy} discovers
equations, which are defined as ``interesting'' if they cannot be simplified
using previously discovered equations. The intuition for such criteria is to
avoid special cases of known theorems, such as $0 + 0 = 0$, $0 + 1 = 1$,
etc. when we already know $\forall x. 0 + x = x$. To work effectively, this
requires general forms to be found \emph{before} special cases.

In the following, we give an overview of the state of the art in automated
theory exploration; identify potential improvements; and present our initial
attempts to implement some of these ideas.

\section{Theory Exploration in Haskell}

Automated theory exploration has been applied to libraries in Isabelle
and Haskell, although we focus on the latter as its implementations are
the most mature (demonstrated by the fact that \textsc{Hipster} explores
Isabelle by first translating it to Haskell). Haskell is interesting to target,
since its use of pure functions and algebraic datatypes causes many functions to
follow algebraic laws; however, since Haskell's type system cannot encode such
laws (without difficulty \cite{lindley2014hasochism}, at least), less
effort is given to finding and stating them; compared to a theorem proving
setting like Isabelle, where these laws are often the main focus.

Due to Haskell's relative popularity, there are also large code repositories
such as \textsc{Hackage} available to explore. This gives several potential
benefits: in the short term, existing library authors and users may be assisted
in comprehending and maintaining their code \cite{QuickSpec}; more notably, TE
could be used to data mine new abstractions, such as typeclasses, with relevance
to real-world code.

Currently, the most powerful TE system for Haskell is \textsc{HipSpec}.
This uses \textsc{QuickSpec} to search through \emph{expressions}
(combinations of the Haskell terms given by the theory), rather than
searching through the space of equations or proofs directly. Expressions
are grouped into equivalence classes, such that the \textsc{QuickCheck}
counterexample finder cannot distinguish between the elements; equations
relating the members of these classes are then conjectured, and sent to
existing automated theorem provers to try and prove \cite{rosen2012proving}. This
approach works well as a lemma generation system, making
\textsc{HipSpec} a capable inductive theorem prover as well as a theory
exploration system \cite{claessen2013automating}.

\section{Identifying Improvements}

Even with a slow search algorithm, we can use a divide and conquor
approach to limit the number of allowed combinations, either by using
stricter types to prevent composition, or by partitioning the theory
into small independently-searched sub-theories. Of course, such
restrictions should strike a balance between the efficiency gained and
the potential to forbid some interesting theorems.

Given this state of the art, we identify the following as potential
areas for improvement:

For a more thorough treatment,
surveys recent work on search algorithms and  surveys
interestingness criteria in the context of data mining.

\begin{itemize}
\item
  Enumerating expressions is a brute-force approach. Identifying more scalable
  alternatives to brute-force search is a well-studied area of Artificial
  Intelligence and Machine Learning. In particular, there is potential in
  applying heuristic algorithms like those surveyed in \cite{blum2011hybrid}.
\item
  Many ``interestingness'' criteria are surveyed in
  \cite{geng2006interestingness}, but existing systems are tightly coupled to
  one particular notion. Whilst pre- and post-processors could provide
  additional filtering, there is merit in building a more loosely-coupled
  alternative, for two reasons:

  \begin{itemize}
  \item
    As the number of irreducible equations grows, it may be desirable to
    impose extra conditions of a more subjective nature.
  \item
    Surprising, insightful equations may be discarded by current systems, if
    they are actually reducible in some complex, non-obvious way. A more
    subjective interestingness measure could be used to veto such
    rejections.
  \end{itemize}
\item
  HipSpec does not propose candidate equations by data mining
  previous results; generalisation methods like anti-unification could
  do this, and at the same time remove the requirement that general
  forms must be enumerated early.
\item
  All type-safe combinations of the given expressions are tried, whilst
  it may be discernable a priori that some combinations are not worth
  considering (either because they are never related, or because their
  relations are never interesting).
\end{itemize}

We consider all of these worth investigating, although in this paper we
focus on the last item.

\section{The \textsc{ML4HS} Framework}

We are implementing a system called \textsc{ML4HS} to investigate scalable
automated theory exploration for Haskell. At this stage, we provide a
pre-processor for \textsc{QuickSpec} which attempts to make exploration of large
theories more tractable by splitting them into a collection of smaller theories.

Inspired by premise selection in automated theorem proving
\cite{kuhlwein2012overview}, and similarity-based clustering method such as that
of \textsc{ML4PG} \cite{journals/corr/abs-1302-6421}, we are investigating the
hypothesis that similar expressions are more likely to be related by equational
properties than dissimilar expressions. This has lead to a divide and conquor
approach, shown in figure 1.

Since our aim is to scale up theory exploration, our pre-processor also
increases the level of automation compared to existing systems. We consider
\textsc{Hackage} packages to be theories, and treat package management,
downloading, compiling, dependency resolution, etc. as necessary tasks for a TE
system such as \textsc{ML4HS} to implement.

Not only does this eliminate the need to define theories manually, as a sub-set
of available expressions, but may also be useful in its own right as a mechanism
to execute arbitrary Haskell code from arbitrary modules in arbitrary packages.

% Acknowledgements for colleagues, referees, ...
% ==============================================
\begin{acknowledge}
I am grateful for those who have helped formulate these ideas through
conversation, especially the HipSpec team at Chalmers University (Moa
Johansson, Koen Claessen, Nick Smallbone, Dan Ros{\'e}n and Irene Lobo Valbuena)
and my supervisor Katya Komendantskaya. I also wish to thank the implementors
of the systems we are building on, including HipSpec, QuickSpec and
QuickCheck on the theory exploration side, as well as GHC, Cabal and Nix
on the infrastructure side.
\end{acknowledge}

% Bibliography with BibTeX
% ========================
\bibliographystyle{eceasst}
\bibliography{../Bibtex}

\end{document}