### 2015-10-29 PhDMeetings/2015-09-16/summer.tex
Additions

\documentclass{article}

\begin{filecontents*}{\jobname.64}
     /-------\
     |Hackage|
     \-------/
         |
         V
   +----------+
   |AST Plugin|
   +----------+
         |    /----\
         |    |    |
         V    V    |
   +------------+  |
   |AST database|  | annotate
   +------------+  |
         |    |    |
         V    \----/
   +------------+
   |Monomorphise|
   +------------+
         |
         V
 +-----------------+
 |Theory definition|
 +-----------------+
         |
         V
+------------------+
|Theory exploration|
+------------------+
\end{filecontents*}

\usepackage{graphicx}
\begin{document}
\title{Summer 2015 Report}

\section{Theory Exploration System}\label{theory-exploration-system}

A lot of my time over summer was spent on building/improving/fixing the wrapper
around QuickSpec which I've dubbed \texttt{ML4HS}. Whilst theoretically
straightforward, peculiarities of Haskell and its infrastructure (GHC,
Cabal, etc.) has lead to several difficulties when trying to build
QuickSpec signatures. The common theme is that it would be useful to
have a function like
\texttt{eval :: [PackageName] -> [ModuleName] -> Expr -> IO (Maybe Expr)}
to (attempt to) evaluate Haskell code in the context of other packages
and modules. This way, we could simply ignore values which lead to
errors.

Until someone implements such a thing, our QuickSpec integration
is instead rather fragile: we must identify and remove any potentially
troublesome values long before they reach QuickSpec, since most of the
errors we encounter are unrecoverable:

\begin{itemize}
\item
  Syntax errors. These can occur when extracting functions from Core,
  since implicit values like type parameters and typeclass dictionaries
  are reified to values with invalid Haskell names (eg.
  \texttt{\$c\textless{}} for an implementation of
  \texttt{<})
\item
  Type errors. These can occur when we try to add functions to a
  signature which do not implement the required typeclasses
  (\texttt{Ord} and \texttt{Typeable}).
\item
  Scope errors. These can occur if we try to use values which are not
  exported.
\item
  GHC panics. These can occur due to mishandling of GHC's API.
\item
  All of the above, at the Template Haskell level (which we use for
  monomorphising)
\end{itemize}

An extensive test suite for every compoenent has helped pin-down and eliminate
most of the sources of errors. The one known bug is that we cannot look up the
names of some packages during compilation: we have their \emph{package key}, but
this is a hash referring to a particular compiled copy on disk; we would rather
have the Hackage name, so we can identify dependencies across packages and
across compilations. The function we currently use to look up names causes a GHC
panic for some packages, although this hasn't been investigated yet.

Conceptually, the project is quite straightforward:

\immediate\write18{ditaa \jobname.64 \jobname-tmp.png}
\fbox{\includegraphics[width=6cm]{\jobname-tmp.png}}

We use a custom GHC plugin to render Core ASTs for top-level values as they are
encountered. These ASTs are labelled with the value's name, its module name and
its package name. This triple acts as a unique ID within the system. Inside the
AST, any references to variables, data constructors or type constructors are
rendered as s-expressions denoting the ID being referenced.

These labelled ASTs are then fed through a series of \emph{annotation} scripts.
Each one appends extra labels to the ASTs, including:

\begin{itemize}
  \item
    The value's type
  \item
    The value's arity (between 0 and 5)
  \item
    The value's \emph{dependencies}, ie. the set of IDs it references.
  \item
    The value's \emph{feature vector}, as determined by one of several implementations.
  \item
    The value's \emph{cluster}, as determined by Weka based on the feature vectors
\end{itemize}

The dependency annotation was implemented by Ouanis, who has also made a simple
script for topologically sorting the ASTs based on their dependencies.

Currently, the \texttt{ML4HSFE} feature extraction script performs a similar
AST-to-matrix transformation as ML4PG, replacing IDs by their cluster as given
by the AST database. If an ID is not found, a default value (eg. ``0'') is used.

To implement recurrent feature extraction, we simply need to perform Ouanis's
sort operation on the database before performing feature extraction, and add extra
calls to the clusterer after each simply-connected-component (ie. mutually-recursive set of functions) has been processed.

\subsection{Going Forward}

The \texttt{ML4HSFE} feature extractor is meant to emulate that of ML4PG (representing the AST nodes in a matrix). We can also use a feature extractor similar to the ACL2 approach (classifying AST nodes based on arities, etc. and using that as the matrix), or similar to Sledgehammer and the ATP work of Josef Urban (counting occurrences of each symbol; I already have a prototype of this).

Moa is also interested in using the output of these tools, so I think it makes sense to release these, eg. to Hackage, sooner rather than later (they are already available from \texttt{chriswarbo.net/git} but may be difficult to install, as Ouanis found).

\section{Reading}\label{reading}

Over the summer I've read many papers, mostly in machine learning, to
better understand the landscape of approaches, their relationships,
history and the current state of the art.

I've also started following Andrew Ng's machine learning lectures from
Stanford.

I've particularly focused on k-means and associated algorithms,
specifically to understand the following:

\begin{itemize}
\item
  How repeatable are its results? \cite{bottou1994convergence}
\item
  How to choose parameters, eg. the number of clusters
  \cite{pelleg2000x} and their initial locations \cite{arthur2007k}?
\item
  How to ensure scalability/tractability? \cite{bahmani2012scalable}
\end{itemize}

I've also kept a few questions in mind regarding recurrent clustering
(RC):

\begin{itemize}
\item
  Can we guarantee convergence of RC?
\item
  How repeatable/replicable is RC (eg. do results vary widely based on
  initial conditions)?
\item
  How does RC compare with in-lining of definitions?
\end{itemize}

As well as AI/ML methods, I've also looked into their application to
Mathematics. From early theorem provers like the Geometry Theorem Prover (which uses counterexamples to eliminate obvious non-theorems before trying to prove them), to exploratory
applications in universal algebra \cite{spector2008genetic}. Other
reading has been ad-hoc, along the lines of either following citation
chains or proceedings (eg. I've been working my way through the
Journal of Machine Learning Research).

\bibliographystyle{plain}
\bibliography{Bibtex}

\end{document}