### 2015-10-29 PhDMeetings/2015-09-16/test.tex

\documentclass{article}
\begin{filecontents*}{\jobname.64}
JVBERi0xLjUKJbXtrvsKMyAwIG9iago8PCAvTGVuZ3RoIDQgMCBSCiAgIC9GaWx0ZXIgL0ZsYXRl
RGVjb2RlCj4+CnN0cmVhbQp4nJWTO1LEMAyG+5xCNYXQy69j7BFoWIpsAdx/Bjlksx4nZE0VWSP/
3y9Z+ZwIY46SAjRBThRyhq8rvL4RXL8nAlFDTQw3jxTZvAxmkCKohSEKclzOaCF5SV6KZ/iY3l8c
IZSyKTSBFmGvdASBkqFZce0H50BrTdTqsNBW8cKapNpeA66qXfFts6rEGKP9w7zfKJGhCaJokNKb
vw9GyStcaD5kHvfgQ3HzsAW/PXSCz2bfw1btIMH9Pb7JCsmA+YbVvsvZ1DmonzMYJeRUx86p+F4Z
GGfMXHXvGRXBEkPNrLdUGP0zwzSwNuxGPTK/k6nu5ZpQUTSRI9ne3dAL7zkVoGCaUIo2DTUt/gU6
+9U6/xtIJSzkocmdrVNnfAyw6+3ZUg3Ma4ftX25hXKYfzwLt9wplbmRzdHJlYW0KZW5kb2JqCjQg
MCBvYmoKICAgMzA1CmVuZG9iagoyIDAgb2JqCjw8CiAgIC9FeHRHU3RhdGUgPDwKICAgICAgL2Ew
IDw8IC9DQSAxIC9jYSAxID4+CiAgID4+Cj4+CmVuZG9iago1IDAgb2JqCjw8IC9UeXBlIC9QYWdl
CiAgIC9QYXJlbnQgMSAwIFIKICAgL01lZGlhQm94IFsgMCAwIDI5Mi4zODk5MjMgNDM3LjI5MzI0
MyBdCiAgIC9Db250ZW50cyAzIDAgUgogICAvR3JvdXAgPDwKICAgICAgL1R5cGUgL0dyb3VwCiAg
ICAgIC9TIC9UcmFuc3BhcmVuY3kKICAgICAgL0NTIC9EZXZpY2VSR0IKICAgPj4KICAgL1Jlc291
cmNlcyAyIDAgUgo+PgplbmRvYmoKMSAwIG9iago8PCAvVHlwZSAvUGFnZXMKICAgL0tpZHMgWyA1
IDAgUiBdCiAgIC9Db3VudCAxCj4+CmVuZG9iago2IDAgb2JqCjw8IC9DcmVhdG9yIChjYWlybyAx
LjExLjIgKGh0dHA6Ly9jYWlyb2dyYXBoaWNzLm9yZykpCiAgIC9Qcm9kdWNlciAoY2Fpcm8gMS4x
MS4yIChodHRwOi8vY2Fpcm9ncmFwaGljcy5vcmcpKQo+PgplbmRvYmoKNyAwIG9iago8PCAvVHlw
ZSAvQ2F0YWxvZwogICAvUGFnZXMgMSAwIFIKPj4KZW5kb2JqCnhyZWYKMCA4CjAwMDAwMDAwMDAg
NjU1MzUgZiAKMDAwMDAwMDcwNSAwMDAwMCBuIAowMDAwMDAwNDE5IDAwMDAwIG4gCjAwMDAwMDAw
MTUgMDAwMDAgbiAKMDAwMDAwMDM5NyAwMDAwMCBuIAowMDAwMDAwNDkxIDAwMDAwIG4gCjAwMDAw
MDA3NzAgMDAwMDAgbiAKMDAwMDAwMDg5NyAwMDAwMCBuIAp0cmFpbGVyCjw8IC9TaXplIDgKICAg
L1Jvb3QgNyAwIFIKICAgL0luZm8gNiAwIFIKPj4Kc3RhcnR4cmVmCjk0OQolJUVPRgo=
\end{filecontents*}
\usepackage{graphicx}

\begin{document}

\immediate\write18{base64 -d \jobname.64 > \jobname-tmp.pdf}

picture is

\fbox{\includegraphics[width=3cm]{\jobname-tmp.pdf}}

\end{document}