### 2015-11-12 TransferReport/report.tex

\documentclass[]{article}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage[T1]{fontenc}
%\usepackage{enumitem}
\usepackage{paralist}
\usepackage{csquotes}
\usepackage[affil-it]{authblk}
\usepackage{listings}
\usepackage{fixltx2e}
\usepackage[numbers]{natbib}
\usepackage{mathpartir}
\usepackage{mmm}

% Treat paragraph as subsubsubsection
\usepackage{titlesec}
\setcounter{secnumdepth}{4}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,decorations.pathreplacing,calc}

\DeclareMathOperator{\md5}{MD5}

\lstset{
  frame=none,
  xleftmargin=2pt,
  stepnumber=1,
  numbers=none,
  numbersep=5pt,
  numberstyle=\ttfamily\tiny\color[gray]{0.3},
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  escapeinside={*'}{'*},
  language=haskell,
  tabsize=2,
  emphstyle={\bf},
  commentstyle=\it,
  stringstyle=\mdseries\rmfamily,
  showspaces=false,
  keywordstyle=\bfseries\rmfamily,
  columns=flexible,
  basicstyle=\small\sffamily,
  showstringspaces=false,
  morecomment=[l]\%,
}

\newcommand{\fc}{System F\textsubscript{C}}

\newcommand{\qcheck}{\textsc{QuickCheck}}
\newcommand{\qspec}{\textsc{QuickSpec}}
\newcommand{\hspec}{\textsc{HipSpec}}

\providecommand{\hs}[1]{\lstinline[language=Haskell]|#1|}

\newenvironment{haskell}
  {
    \begin{lstlisting}[language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
  }
  {
    \end{lstlisting}
  }

\providecommand{\coq}[1]{\lstinline[language=ML]|#1|}

\begin{document}

\pagestyle{headings}  % switches on printing of running heads

\title{Interestingness Measures for Theory Exploration}

\author{Chris Warburton}

\affil{University of Dundee,\\
\texttt{http://tocai.computing.dundee.ac.uk}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Since their inception, computers have been applied to automate and assist work in mathematics, with established fields including numerical computation, computer algebra, (online) communication and so on. Here we focus on their use in two areas: formal proof and statistics; and how new approaches are bringing these seemingly disparate disciplines closer together.
\end{abstract}

\input{intro.tex}
\input{background.tex}
\input{examples.tex}
\input{contributions.tex}
\input{implementation.tex}

\section{First Results}

\iffalse
TODO

Future work?
Hypothetical use cases?
\fi

\input{related.tex}

\section{Future Work}
\label{future}

\iffalse
TODO

interestingness
\fi

\iffalse

QuickSpec: extend or extinguish?

Improve and find other use cases/scenarios for clustering and feature extraction

Other directions for Theory Exploration?

How about systems based on term rewriting, logic programming, etc.?
\fi

\section{Conclusion}
\label{conclusion}

\bibliographystyle{plain}
\bibliography{../Bibtex}

\end{document}