### 2015-11-17 TransferReport/report.tex

\documentclass[]{article}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{algorithm2e}
\usepackage[T1]{fontenc}
%\usepackage{enumitem}
\usepackage{paralist}
\usepackage{csquotes}
\usepackage[affil-it]{authblk}
\usepackage{listings}
\usepackage{fixltx2e}
\usepackage[numbers]{natbib}
\usepackage{mathpartir}
\usepackage{mmm}

% Treat paragraph as subsubsubsection
\usepackage{titlesec}
\setcounter{secnumdepth}{4}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,decorations.pathreplacing,calc}

\DeclareMathOperator{\md5}{MD5}

\lstset{
  frame=none,
  xleftmargin=2pt,
  stepnumber=1,
  numbers=none,
  numbersep=5pt,
  numberstyle=\ttfamily\tiny\color[gray]{0.3},
  columns=flexible,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  escapeinside={*'}{'*},
  language=haskell,
  tabsize=2,
  emphstyle={\bf},
  commentstyle=\it,
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keepspaces=true,
  keywordstyle=\bfseries\ttfamily,
  basicstyle=\small\ttfamily,
  showstringspaces=false,
  morecomment=[l]\%,
}

\newcommand{\fc}{System F\textsubscript{C}}

\newcommand{\qcheck}{\textsc{QuickCheck}}
\newcommand{\qspec}{\textsc{QuickSpec}}
\newcommand{\hspec}{\textsc{HipSpec}}

\providecommand{\hs}[1]{\lstinline[language=Haskell]|#1|}

\providecommand{\coq}[1]{\lstinline[language=ML]|#1|}

\lstnewenvironment{haskell}{%
  \lstset{language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth}}{}

\lstnewenvironment{coqblock}{%
  \lstset{language=ML, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth}}{}

\begin{document}

\pagestyle{headings}  % switches on printing of running heads

\title{Improving Haskell Theory Exploration}

\author{Chris Warburton}

\affil{University of Dundee,\\
\texttt{http://tocai.computing.dundee.ac.uk}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Theory Exploration is a promising approach to improving the quality and understanding of software, above that available through testing, in languages which are amenable to formal analysis such as those based on pure functional programming. Current techniques are limited by their computational cost, which we mitigate through a divide and conquer approach: reducing the problem size whilst using a machine learning approach to preserve relevant details.
\end{abstract}

\input{intro.tex}
\input{background.tex}
\input{contributions.tex}
\input{implementation.tex}

\iffalse
\section{First Results}

TODO

Future work?
Hypothetical use cases?
\fi

\input{related.tex}

\section{Future Work}
\label{future}

\iffalse
TODO

interestingness
\fi

\section{Conclusion}
\label{conclusion}

\bibliographystyle{plain}
\bibliography{../Bibtex}

\end{document}