### 2015-11-24 TransferReport/evaluation.tex

\iffalse
\section{First Results}

TODO

Future work?
Hypothetical use cases?
\fi