### 2015-11-30 TransferReport/evaluation.tex
Added TODOs

\section{Evaluation}

TODO

Future work?
Hypothetical use cases?