### 2015-12-09 TransferReport/report.tex

\documentclass[]{article}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[T1]{fontenc}
%\usepackage{enumitem}
\usepackage{paralist}
\usepackage{csquotes}
\usepackage[affil-it]{authblk}
\usepackage{listings}
\usepackage{fixltx2e}
\usepackage[numbers]{natbib}
\usepackage{mathpartir}
\usepackage{mmm}
\usepackage{subcaption}
\usepackage{tikz-qtree}

% Treat paragraph as subsubsubsection
\usepackage{titlesec}
\setcounter{secnumdepth}{4}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,decorations.pathreplacing,calc}

\DeclareMathOperator{\md5}{MD5}

\lstset{
  frame=none,
  xleftmargin=2pt,
  stepnumber=1,
  numbers=none,
  numbersep=5pt,
  numberstyle=\ttfamily\tiny\color[gray]{0.3},
  columns=flexible,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  escapeinside={*'}{'*},
  language=haskell,
  tabsize=2,
  emphstyle={\bf},
  commentstyle=\it,
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keepspaces=true,
  keywordstyle=\bfseries\ttfamily,
  basicstyle=\small\ttfamily,
  showstringspaces=false,
  morecomment=[l]\%,
}

\newcommand{\fc}{System F\textsubscript{C}}

\newcommand{\blank}{\cdot}
\newcommand*\mean[1]{\overline{#1}}
\newcommand*\vect[1]{\mathbf{#1}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\feature}[1]{\left\ulcorner #1 \right\urcorner}
\newcommand{\id}[1]{\texttt{"#1"}}
\newcommand{\vlocal}[1]{\CVar\ (\CLocal\ \id{#1})}
\newcommand{\vglobal}[1]{\CVar\ (\CGlobal\ \id{#1})}
\newcommand{\cat}{\mbox{\ensuremath{+\!\!+\,}}}

\newcommand{\qcheck}{\textsc{QuickCheck}}
\newcommand{\qspec}{\textsc{QuickSpec}}
\newcommand{\hspec}{\textsc{HipSpec}}
\newcommand{\equal}{=}
\newcommand{\hs}[1]{\texttt{#1}}

\providecommand{\coq}[1]{\lstinline[language=ML]|#1|}

\lstnewenvironment{haskell}{%
  \lstset{language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth}}{}

\lstnewenvironment{coqblock}{%
  \lstset{language=ML, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth}}{}

\begin{document}

\pagestyle{headings}  % switches on printing of running heads

\title{Improving Haskell Theory Exploration}

\author{Chris Warburton}

\affil{University of Dundee,\\
\texttt{http://tocai.computing.dundee.ac.uk}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Theory Exploration is a promising approach to improving the quality and understanding of software. It extends previously existing methods available through testing, in languages which are amenable to formal analysis such as those based on pure functional programming. Current theory exploration techniques are limited by their use of exponential time algorithms, which we mitigate by intelligently reducing the size of large inputs, using machine learning to preserve the relationships needed for successful exploration.

\iffalse TODO: I would emphasise more selective (intelligent) approach, and as a result - better precision of TE. Rather than saying that it is merely a size/complexity problem. \fi
\end{abstract}

\input{intro.tex}
\input{background.tex}
\input{contributions.tex}
\input{implementation.tex}
\input{related.tex}
\input{evaluation.tex}
\input{future.tex}
\iffalse TODO: Section that brings back ML discussion to TE is missing \fi

\bibliographystyle{plain}
\bibliography{../Bibtex}

\end{document}