### 2017-02-15 benchmark2016/article.tex
Add benchmark2016

\title{An Approach to Benchmark Conjecture Generators}
\author{
        Chris Warburton \\
                University of Dundee\\
}
\date{\today}

\documentclass[12pt]{article}

\begin{document}
\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

Many algorithms and approaches have been developed to automate the discovery of new mathematical theorems, but it is difficult to compare their performance in a quantitative way. Existing approaches include precision/recall measurements on existing theories, but these rely on some form of ground truth, i.e. which theorems are ``worth'' producing.

We propose that the extensive test suites used to benchmark automated theorem provers can be repurposed to create benchmarks for theory exploration systems. We provide a tool to automatically perform this conversion, given an existing benchmark suite in SMT-LIB format \cite{BarFT-SMTLIB}.

\section{Previous approaches}\label{previous approaches}

Theory exploration tools such as IsaCoSy \cite{Johansson.Dixon.Bundy:conjecture-generation}, QuickSpec and IsaScheme have been evaluated using the standard library of the Isabelle theorem prover as their ground truth.

ALGORITHM:

Concatenate all definitions from all files together.
For each definition, prefix its name with the filename it came from and update all references.
Normalise all local variable names.
Remove any identical definitions, leaving the first one in place and updating any references to the removed definitions to use this one.
Repeat, until no further removals are possible.

\bibliographystyle{plain}
\bibliography{./Bibtex}

\end{document}