### 2018-04-06 Bucketing/intro.tex
Bucketing changes (mostly long lines)

\section{Introduction}

Functional programming claims to be more ``mathematically tractable'' than other
``conventional'' approaches, due to properties like \emph{referential
transparency} and \emph{immutability}, which allow equational reasoning without
many of the complications caused by side effects and strict evaluation
order~\cite{hughes1989functional}. Languages such as Haskell combine these
advantages with static type systems to restrict the possible values of every
expression, making programs amenable to powerful analysis tools such as property
checking~\cite{claessen2011quickcheck} and theorem
proving~\cite{rosen2012proving}.

Whilst such tools can support or refute user-provided properties, we are
interested in the related problem of \emph{discovering} novel properties
automatically; a task addressed by the field of \emph{Mathematical Theory
  Exploration}. Whilst clever algorithms are used to guide their search,
existing MTE tools, including those applied to Haskell, ultimately rely on
enumerating all (well-typed) expressions. This limits their usefulness, as it is
only feasible to explore a small collection of functions at a time, relying on
the user to select promising combinations. We investigate the use of machine
learning methods to perform this selection automatically, removing the need for
user intervention whilst maintaining the ability to discover desirable
properties.

Our contributions are:

\begin{enumerate}
  \item The application of machine learning algorithms to theory exploration,
    for intelligently discovering interesting sub-sets of Haskell libraries,
    which are more tractable to explore.
  \item A novel feature extraction method for transforming Haskell expressions
    into a form amenable to off-the-shelf learning algorithms.
  \item An implementation of these feature extraction and theory exploration
    approaches.
  \item A comparison of our methods with existing approaches, both for theory
    exploration in Haskell, and for machine learning in other languages.
\end{enumerate}

We describe our treatment of Haskell expressions in \S~\ref{sec:haskell} and
describe the \qspec{} theory exploration system we build upon in
\S~\ref{sec:theoryexploration}. We discuss our contributions in more depth in
\S~\ref{sec:contributions}, and provide implementation details in
\S~\ref{sec:implementation}. A variety of related work is surveyed in
\S~\ref{sec:related}, we briefly evaluate our implementations in
\S~\ref{sec:evaluation} and give several potential directions for future
research in \S~\ref{sec:future}.