### 2018-04-19 PhDSymposium2018/abstract.tex
Finish symposium text

\documentclass[]{default}
\usepackage{lmodern}
\usepackage{listings}
\usepackage{amssymb,amsmath}
\usepackage{paralist}
\usepackage{tikz}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Quantitative Benchmarking for Conjecture Generation},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}
\usepackage{subcaption}
\newcommand{\etal}{{\em et al.}}

\title{Quantitative Benchmarking for Conjecture Generation}
  \author{
          Chris Warburton \\
      University of Dundee \\
      \texttt{cmwarburton@dundee.ac.uk}
    }
\date{}

\begin{document}
\maketitle
\begin{abstract}
  We propose a benchmarking methodology to evaluate the efficiency and quality
  of \emph{automated conjecture generation}. Our approach uses widely available
  theorem proving tasks as a \emph{ground-truth} corpus, and we demonstrate its
  use on the \textsc{QuickSpec} and \textsc{IsaCoSy} tools, finding that the
  former takes significantly less time to produce significantly more
  ``interesting'' output. By providing a standard, cross-tool evaluation
  technique we allow the disparate approaches currently being pursued to be more
  directly compared. Our hope is to encourage innovation and improvements
  similar to those seen in fields like automated theorem proving, where the
  availability of benchmarks encourages healthy competition.
\end{abstract}

\section{Introduction}\label{introduction}

\emph{Conjecture generation} (CG) is the open-ended task of conjecturing
``interesting'' properties of a given logical theory (such as a software
library), and is studied as a sub-field of \emph{mathematical theory
  exploration}. Such conjectures can be sent to automated theorem provers (ATP),
have helped prove software correctness by finding key
lemmas~\cite{Claessen.Johansson.Rosen.ea:2013}. Automating formal methods this
way can reduce the cost and difficulty which have historically limited their use
to mission- or life-critical domains. More practical automated reasoning shows
promise in software
engineering~\cite{McKinna:2006,chlipala2011certified,Xi2003}, automated research
``assistants''~\cite{McCarthy_Programs59,lenat:77,benzmuller1997omegamega} and
educational software~\cite{conf/ijcai/TrybulecB85,hendriks2010teaching}.

CG progress is hard to ascertain since there is no standard measure of
``success'', partially due to the ambiguity of the task, i.e. what counts as
``interesting''? Researchers may have different goals, and a variety of
evaluation methods appear in the literature. We attempt to solve this
discrepancy (in the short-term) by defining a standard, unambiguous, extensible
benchmarking approach for CG tools. Our contributions include:

\begin{itemize}
\item A benchmarking methodology for CG tools.
\item Resolving the ``interestingness'' question by adapting ATP benchmarks into
  ground-truth input/output examples.
\item Benchmarking and comparing the \textsc{QuickSpec} and \textsc{IsaCoSy}
  tools.
\end{itemize}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{quickspectime.pgf}
    \caption{\textsc{QuickSpec}}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{isacosytime.pgf}
    \caption{\textsc{IsaCoSy}}
  \end{subfigure}

  \caption{Time taken for CG on samples from TIP, with 300 second timeout. Lines
    show medians and error bars show inter-quartile range, with 31 samples of
    each size except 1 (14) and 2 (30). }
  \label{figure:quickspec_runtimes}
\end{figure}

\begin{figure*}
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{quickspecprec.pgf}
    \caption{\textsc{QuickSpec}}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \input{isacosyprec.pgf}
    \caption{\textsc{IsaCoSy}}
  \end{subfigure}

  \caption{Precision and recall of successful runs. Lines show average
    proportion, with sample standard deviation shaded.}
  \label{figure:quickspec_precRec}
\end{figure*}

\section{Background}

Consider a software library for inductive (i.e. singly-linked) lists, containing
simple functions like \texttt{append}, more complex ones like \texttt{mergesort}
and even higher-order functions like \texttt{map} and \texttt{reduce}. The large
input space of these functions limits \emph{testing} to a tiny fraction of the
possible executions, whilst higher-order functions and induction makes
\emph{proving} correctness properties undecidable (requiring significant user
guidance). We may also \emph{miss} key properties (especially if the same person
writes the spec and the code), e.g. that calling \texttt{mergesort} on something
already sorted can be optimised away. Similar tasks arise in mathematics, and
scientific models more generally, when exploring the consequences of axioms and
definitions.

The aim of CG is to read such definitions and produce a set of ``interesting''
properties which they appear to satisfy. This can help formalise the
``background knowledge'' required for proving, and present the user with ideas
they hadn't considered.

Crucial to CG is how we define ``interesting''. The literature gives many
alternatives, for example Colton \etal{} survey existing
tools~\cite{colton2000notion} and find six common qualities: empirical
plausibility, novelty, surprisingness, applicability, comprehensibility and
utility.

\emph{Quantitative} evaluation has compared the
\textsc{IsaCoSy}~\cite{Johansson.Dixon.Bundy:conjecture-generation},
\textsc{IsaScheme}~\cite{Montano-Rivas.McCasland.Dixon.ea:2012} and
\textsc{HipSpec}~\cite{claessen2013automating} tools, by selecting sets of
definitions and properties from the standard library of the \textsc{Isabelle}
proof assistant, and treating those as a \emph{ground-truth} corpus, i.e. using
``membership in the corpus'' as a proxy for ``interesting''. The proportion of
conjectures appearing in the corpus is the \emph{precision}, the proportion of
the corpus appearing in the conjectures is the \emph{recall}.

This is a pragmatic solution to the ambiguity, but only two datapoints were
taken: one with 12 properties, the other with 9. This makes it hard to draw
robust conclusions.

\section{Methodology}

Building on existing precision/recall work, we define a benchmarking methodology
with a much larger ground-truth corpus, and provide a scalable, statistical
approach to evaluation using this corpus. Inspired by the availability of large
benchmarks for ATP tools~\cite{pelletier2002development}, we take those problem
sets and automatically convert them into input/output examples for the CG task.

We choose an ATP benchmark relevant to our domain. For our demonstration we
chose Tons of Inductive Problems (TIP)~\cite{claessen2015tip}, which is relevant
to software verification due to using higher-order logic, inductive definitions
and relevant problems (arithmetic, sorting, etc.). TIP 0.2 contains 343 theorem
proving problems, each defining the types and functions involved and stating a
single property of those definitions. A total of 618 datatypes and 1498
functions are defined, but most of these are duplicates (e.g. every problem
involving \texttt{multiply} will include its own definitions of
\texttt{multiply}, \texttt{add}, \texttt{successor}, \texttt{Natural}, etc.).
Removing these redundancies, normalising the definitions and only selecting
those which appear in at least one property leaves 182 functions for use in our
corpus.

We could, in theory, send these de-duplicated definitions straight into a CG
tool and use the 343 properties as the ground truth for analysis. However, this
number of definitions is too large to be feasible for current tools, and we
would only get a single datapoint this way. Instead we \emph{sample} a subset of
definitions, and form the ground truth by selecting those properties which only
reference the sampled definitions. To avoid undefined precisions, we only use
samples with at least one ground-truth property.

We translate the function definitions into each tool's input language (Isabelle
for \textsc{IsaCoSy} and Haskell for \textsc{QuickSpec}), run the tools with a
timeout (5 minutes, based on preliminary experiments) and translate the outputs
into a standard form for comparison with the ground-truths.

\section{Demonstration}

We applied our methodology to two existing CG tools:
\textsc{QuickSpec}~\cite{QuickSpec} (a component of
\textsc{HipSpec}~\cite{Claessen_hipspec:automating} and
\textsc{Hipster}~\cite{Hipster}) enumerates expressions in
Haskell~\cite{marlow2010haskell}, conjecturing that they're all equal unless
proven otherwise by a counter-example finder;
\textsc{IsaCoSy}~\cite{johansson2009isacosy} enumerates equations in
\textsc{Isabelle}~\cite{nipkow2002isabelle}, using a constraint solver to avoid
repeating itself.

Their running times are shown in Figure~\ref{figure:quickspec_runtimes}, for
sample of up to 20 functions. The measurements are paired, and we followed the
Speedup-Test protocol of Touati, Worms and Briais~\cite{touati2013speedup} to
find that \textsc{QuickSpec} was faster for $67\%-98\%$ of sizes (with 95\%
confidence). Only sizes 1 and 2 weren't significantly faster, due to too few
samples. Timeouts bounding the difference make this a conservative estimate.

Precision and recall are shown in Figure~\ref{figure:quickspec_precRec},
although few \textsc{IsaCoSy} runs succeeded. Recalls remain roughly flat, as
the algorithms are monotonic (adding definitions can't prevent conjectures),
whilst precisions decrease due to the buildup of ``uninteresting''
relationships. We pooled all successful samples together and compared: McNemar's
test~\cite{mcnemar1947note} found \textsc{QuickSpec}'s recall significantly
higher ($p = 0.0026$), and Boschloo's test~\cite{lydersen2009recommended} found
no significant difference between the precisions ($p = 0.111$).

\section{Conclusion}

``Interestingness'' remains a difficult philosophical question, but we work
around this to provide a practical benchmark which researchers can use for the
foreseeable future to characterise and, importantly, \emph{compare} their
solutions to the CG task. Although it does not capture every facet of the CG
goal, our methodology builds on existing work and is easily extended to other
corpora, tools and analyses.

We have demonstrated that \textsc{QuickSpec} significantly out-performs
\textsc{IsaCoSy}, and hope that such healthy competition will accelerate future
developments, as has been seen in the field of ATP.

\bibliographystyle{plain}
\bibliography{../Bibtex}

\end{document}