### 2018-05-17 PhDThesis/background.tex
Drag a bunch of stuff into thesis folder

\section{Background}

\input{background_haskell.tex}
\input{background_quickcheck.tex}
\input{background_theoryexploration.tex}
\input{background_conjecturegeneration.tex}
\input{background_machinelearning.tex}
\input{background_recurrentclustering.tex}