### 2018-05-17 PhDThesis/evaluation_existingcode.tex

\section{Evaluation}
\label{sec:evaluation}

We have applied our recurrent clustering algorithm to several scenarios, with
mixed results. A major difficulty in evaluating these clusters is that we have
no ``ground truth'', i.e. there is no objectively correct way to compare
expressions. Instead, we provide a qualitative overview of the more interesting
characteristics.

As a simple example, we clustered (Haskell equivalents of) the running examples
used to present ACL2(ml) \cite{heras2013proof}, shown in Figure
\ref{fig:lisp}. These include tail-recursive and non-tail-recursive
implementations of several functions. We expect those with similar
\emph{structure} to be clustered together, rather than those which implement the
same function. The results are shown in Figure \ref{fig:haskellcluster}, where
we can see the ``\hs{Tail}'' functions clearly distinguished, with little
distinction between the tail recursive and na\"{\i}ve implementations.

Next we tested whether these same functions would be clustered together when
mixed with seemingly-unrelated functions, in this case 207 functions from
Haskell's \hs{text} library. In fact, the \hs{helperFib} and \hs{fibTail}
functions appeared together in a separate cluster from the rest. This was
unexpected, with no obvious semantic connection between these two functions and
the others in their cluster (although most are recursive, due to the nature of
the \hs{text} library).

\begin{figure}
  \begin{haskell}
(defun fact (n)
  (if (zp n) 1 (* n (fact (- n 1)))))

(defun helper-fact (n a)
  (if (zp n) a (helper-fact (- n 1) (* a n))))

(defun fact-tail (n)
  (helper-fact n 1))

(defun power (n)
  (if (zp n) 1 (* 2 (power (- n 1)))))

(defun helper-power (n a)
  (if (zp n) a (helper-power (- n 1) (+ a a))))

(defun power-tail (n)
  (helper-power n 1))

(defun fib (n)
  (if (zp n)
      0
      (if (equal n 1)
          1
          (+ (fib (- n 1)) (fib (- n 2))))))

(defun helper-fib (n j k)
  (if (zp n)
      j
      (if (equal n 1)
          k
          (helper-fib (- n 1) k (+ j k)))))

(defun fib-tail (n)
  (helper-fib n 0 1))
  \end{haskell}
  \caption{Common Lisp functions, both tail-recursive and non-tail-recursive.}
  \label{fig:lisp}
\end{figure}

\begin{figure}
  \begin{haskell}
fact n = if n == 0
            then 1
            else n * fact (n - 1)

helperFact n a = if n == 0
                    then a
                    else helperFact (n - 1) (a * n)

factTail n = helperFact n 1

power n = if n == 0
             then 1
             else 2 * power (n - 1)

helperPower n a = if n == 0
                     then a
                     else helperPower (n - 1) (a + a)

powerTail n = helperPower n 1

fib n = if n == 0
           then 0
           else if n == 1
                   then 1
                   else fib (n - 1) + fib (n - 2)

helperFib n j k = if n == 0
                     then j
                     else if n == 1
                          then k
                          else helperFib (n - 1) k (j + k)

fibTail n = helperFib n 0 1
  \end{haskell}
  \caption{Haskell equivalents of the Common Lisp functions in Figure \ref{fig:lisp}.}
  \label{fig:haskelllisp}
\end{figure}

\begin{figure}
  \centering
    \tikzstyle{block} = [rectangle, draw]
    \tikzstyle{line}  = [draw, -latex']
    \begin{tikzpicture}[node distance=3cm]
      \node [block] (c1) {
        \begin{tikzpicture}[node distance = 1cm]
          \node [block]                    (factTail)  {\hs{factTail}};
          \node [block, below of=factTail] (fibTail)   {\hs{fibTail}};
          \node [block, below of=fibTail]  (powerTail) {\hs{powerTail}};
        \end{tikzpicture}
      };
      \node [block, right of=c1] (c2) {
        \begin{tikzpicture}[node distance = 1cm]
          \node [block]                (fact)       {\hs{fact}};
          \node [block, below of=fact] (helperFact) {\hs{helperFact}};
        \end{tikzpicture}
      };
      \node [block, right of=c2] (c3) {
        \begin{tikzpicture}[node distance = 1cm]
          \node [block]                       (fib)         {\hs{fib}};
          \node [block, below of=fib]         (helperFib)   {\hs{helperFib}};
          \node [block, below of=helperFib]   (helperPower) {\hs{helperPower}};
          \node [block, below of=helperPower] (power)       {\hs{power}};
        \end{tikzpicture}
      };
  \end{tikzpicture}
  \caption{Typical clusters for the functions in Figure \ref{fig:haskelllisp}.}
  \label{fig:haskellcluster}
\end{figure}

We have also applied our recurrent clustering algorithm to a variety of the
most-downloaded packages from \textsc{Hackage} (as of 2015-10-30), including
\texttt{text} (as above), \texttt{pandoc}, \texttt{attoparsec},
\texttt{scientific}, \texttt{yesod-core} and \texttt{blaze-html}. Whilst we
expected functions with a similar purpose to appear together, such as the
various reader and writer functions of \hs{pandoc}, there were always a few
exceptions which became separate for reasons which are still unclear.

When clustering the \hs{yesod} Web framework, the clustering did seem to match
our intuitions, in particular since all 15 of Yesod's MIME type identifiers
appeared in the same cluster.

Whilst recurrent clustering has produced results which merit further
investigation, the application to theory exploration has yet to be tested
empirically. This is due to \qspec{}'s use of \qcheck{}'s \hs{Arbitrary} type
class to generate random values for instantiating variables. Whilst we can
automatically define \qspec{} theories and invoke them with \hs{nix-eval}, not
all types have \hs{Arbitrary} instances; those without cannot be given any
variables in our signature, which severely limits the possible combinations
which can be explored. In many cases, no variables can be included at all,
leaving just equations involving constants. This has so far prevented us from
measuring the direct impact on \qspec{} performance, either directly by
exploring the sub-sets identified through recurrent clustering, or indirectly by
comparing the equations generated by a full brute-force search to our recurrent
clusters: those equations relating terms from different clusters would not be
discovered by our method. It is this ratio of equations found through brute
force to those found after narrowing-down by clusters which is one of our key
objectives to maximise at this stage; until we begin to pursue the
\emph{interestingness} of the properties.

The following less-serious problems were also encountered while applying
\textsc{ML4HS} to \textsc{Hackage} packages:

\begin{itemize}
\item Some packages, such as \texttt{warp} and \texttt{conduits}, get no
  declarations to cluster. This is because they make all of their declarations
  privately, e.g. in ``internal'' modules, then use separate modules to export
  the public declarations. GHC's renaming phase makes all references to such
  exports canonical, by pointing them to the private declarations. This forces
  us to ignore such declarations, as \qspec{} will not be able to access them.

\item Since we do not support type-level entities, we ignore type
  classes. Unfortunately, this also means ignoring any value-level bindings (AKA
  ``methods'') which occur in a type class instance. Instead of being clustered,
  these result in references getting $f_{recursion}$ features. This is
  especially noticable in libraries like \hs{scientific}, where only the
  functions for constructing and destructing numbers in scientific notation are
  clustered; all of the arithmetic is defined in type classes. One difficulty
  with supporting methods is that their namespace in Core is disjoint from that
  of regular Haskell identifiers: a transformation layer would be required,
  along with explicit type annotations to avoid ambiguity.
\end{itemize}

It seems like this recurrent clustering method has promise, although it will
require a more thorough exploration of the parameters to obtain more intuitively
reliable results. These clusters can then be used in several ways to perform
theory exploration; the most na\"{\i}ve way being to explore each cluster as
\qspec{} signature in its own right. \textsc{ML4HS} already provides this
functionality, although the lack of test generators severely limits what can be
discovered.

As alluded to previously, we also have the opportunity to reason by
\emph{analogy}. Similar to the work on ACL2(ml) \cite{heras2013proof}, we could
produce a general ``scheme'' from each equation we find (either through \qspec{}
or by data mining test suites); like Isabelle approaches have shown
\cite{Montano-Rivas.McCasland.Dixon.ea:2012}, these schemes could then be
instantiated to a variety of similar values, in an attempt to find new theorems
which are analogues of existing results from a different context. ``Mutating''
existing theorem statements in such a way would also increase the chance of any
result being considered interesting; since it's likely that the unmodified
statement was deemed interesting, and the new result would not in general follow
as a simple logical consequence.