### 2018-05-17 PhDThesis/litreview_interestingness.tex

Whilst there are many reasonably objective benchmarks for mathematical tasks
such as automated theorem proving, the precision/recall analysis shown
in~\cite{claessen2013automating}, and described further in $\S$\ref{sec:te}, is
the only quantitative comparison of these recent MTE tools we have found in the
literature. Our work is essentially an extension of this approach, to a larger
and more diverse set of examples. The suitability of the TIP theorem proving
benchmark for our purposes, detailed in $\S$\ref{sec:tip}, is not coincidental,
since its developers are also those of the IsaCoSy and QuickSpec tools we have
tested. This goes some way to ensuring that our demonstration is a faithful
representation of the task these tools were intended to solve; our independent
repurposing of this problem set, in a way it was not designed for, reduces the
risk that the benchmark is tailor-made for these tools (or that the tools
over-fit to these particular problems).