### 2018-05-17 PhDThesis/litreview_theoryformation.tex

An important historical implementation of ATF is Lenat's AM (Automated
Mathematician) system. Unlike prior work, such as
Meta-Dendral~\cite{buchanan:75} and those described in~\cite{winston}, AM aims
to be a general purpose mathematical discovery system, designed to both
construct new concepts and conjecture relationships between them. AM is a
rule-based system which represents knowledge using a frame-like scheme, enlarges
its knowledge base via a collection of heuristic rules, and controls the firing
of these rules via an agenda mechanism. Evaluation of AM considered generality
(performance in new domains) and how finely-tuned various aspects of the program
are (the agenda, the interaction of the heuristics, etc). Most of this
evaluation was qualitative, and has subsequently been
criticised~\cite[chap.~13]{colton:book}. In their case study in methodology,
Ritchie and Hanna found a large discrepancy between the theoretical claims made
of AM and the implemented program~\cite{ritchie1984case}; for example, AM ``invented''
natural numbers from sets, but did so using a heuristic specifically designed to
make this connection.