### 2018-05-17 PhDThesis/litreview_premiseselection.tex
Reduce some duplication

\subsubsection{Relevance Filtering}
\label{sec:relevance}

% TODO
\cite{kuhlwein2012overview}

The combinatorial nature of formal systems causes many proof search methods,
such as resolution, to have exponential complexity
\cite{haken1985intractability}; hence even a modest size increase can turn a
trivial problem into an intractable one. Finding efficient alternatives for such
algorithms, especially those which are NP-complete (e.g. determining
satisfiability) or co-NP-complete (e.g. determining tautologies), seems
unlikely, as it would imply progress on the famously intractable open problems
of $\text{P} = \text{NP}$ and $\text{NP} = \text{co-NP}$. On the other hand, we
can turn this difficulty around: a modest \emph{decrease} in size may turn an
intractable problem into a solvable one. We can ensure that the solutions to
these reduced problems coincide with the original if we only remove
\emph{redundant} information. This leads to the idea of \emph{relevance
  filtering}.

Relevance filtering simplifies a proof search problem by removing from
consideration those clauses (axioms, definitions, lemmas, etc.) which are deemed
\emph{irrelevant}. The technique is used in Sledgehammer during its translation
of Isabelle/HOL theories to statements in first order logic: rather than
translating the entire theory, only a sub-set of relevant clauses are
included. This reduces the size of the problem and speeds up the proof search,
but it creates the new problem of determining when a clause is relevant: how do
we know what will be required, before we have the proof?

The initial approach, known as \textsc{MePO} (from \emph{Meng-Paulson}
\cite{meng2009lightweight}), gives each clause a score based on the proportion
$m / n$ of its symbols which are ``relevant'' (where $n$ is the number of
symbols in the clause and $m$ is the number which are relevant). Initially, the
relevant symbols are those which occur in the goal, but whenever a clause is
found which scores more than a particular threshold, all of its symbols are then
also considered relevant. There are other heuristics applied too, such as
increasing the score of user-provided facts (e.g. given by keywords like
\texttt{using}), locally-scoped facts, first-order facts and rarely-occuring
facts. To choose $r$ relevant clauses for an ATP invocation, we simply order the
clauses by decreasing score and take the first $r$ of them.

Recently, a variety of alternative algorithms have also been investigated,
including:

\begin{description}
\item{\textsc{MaSH}}: Machine Learning for SledgeHammer
  \cite{kuhlwein2013mash}. The distinguishing feature of \textsc{MaSH} is its
  use of ``visibility'', which is essentially a dependency graph of which
  theorems were used in the proofs of which other theorems; although theorems
  are represented as abstract sets of features. To select relevant clauses for a
  goal, the set of clauses which are visible from the goal's components is
  generated; this is further reduced by (an efficient approximation of) a naive
  Bayes algorithm.

\item{\textsc{MOR}}: \emph{Multi-output ranking} uses a support vector machine
  (SVM) approach for selecting relevant axioms from the Mizar Mathematical
  Library for use by the Vampire ATP system \cite{alama2014premise}. \iffalse
  TODO: describe the kernel, as that's the interesting bit \fi It compares
  favourably to \textsc{SNoW} and \textsc{SInE}.

  % TODO:
\item{\textsc{SInE}}
\item{\textsc{BliStr}}
\item{\textsc{HOLyHammer}}
\item{\textsc{MoMM}}
\item{\textsc{SNoW}}
\item{\textsc{MPTP 0.2}}
\item{\textsc{MaLARea}}
\item{\textsc{MaLARea SG1}}

\end{description}