### 2018-09-12 PhDThesis/litreview.tex
Thesis changes (e.g. lit review)

\section{Literature Review}

\input{litreview_automatedreasoning.tex}
\input{litreview_staticanalysis.tex}
\input{litreview_testing.tex}
\input{litreview_premiseselection.tex}
\input{litreview_interestingness.tex}