### 2018-09-12 PhDThesis/litreview_machinelearning.tex

\subsection{Machine Learning Applications in Mathematics, Logic and Programming Language Theory}

Machine learning techniques, used for finding statistical patterns and numerical
correlations in arbitrary data, are characterised by their statistical pattern-finding

AutoTuning

DeepTune analyses raw program text to predict the best options for an optimising
compiler. FIXME

Fuzzing AFL DeepSmith