### 2018-09-13 PhDThesis/background.tex
Thesis bits

\section{Background}

% Background should cover approaches to automated reasoning about software, and
% formal systems in general. We can highlight how the language/representation
% imposes invariants and guarantees on what we can know, and how we can extend
% those guarantees with features like purity (prevents behaviour depending on
% ambient environment/state), lazy evaluation (prevents problems with
% non-termination), sound type systems (limits the space of possible values for
% each expression), etc. and therefore how Haskell is a natural vehicle for us to
% use. Giving background on Haskell naturally leads to descendants like Agda and
% Idris, which are also interactive theorem provers; which lets us discuss
% Isabelle, Coq, HOL, etc., how they're difficult for existing reasoning
% algorithms to handle (e.g. resolution, which works well in first-order logics),
% approaches to helping

\input{background_haskell.tex}
\input{background_quickcheck.tex}
\input{background_theoryexploration.tex}
\input{background_conjecturegeneration.tex}
\input{background_machinelearning.tex}
\input{background_recurrentclustering.tex}