### 2019-02-28 UnderstandabilityPaper/machinelearning.tex
Fleshed out automated theorem proving section

% STRUCTURE:

%  - Understandability in AI
%  - Theorem proving as example:
%   - Least informative is a yes/no/unknown answer
%   -

\documentclass{article}
\begin{document}
The problem of making machine-generated artefacts more understandable to human
users is quite clear in the field of automated theorem proving~CITE: the task of
deciding whether a given formal statement follows from a given set of premeses.
The least informative approaches, such as those based on the popular \emph{SLD
  resolution} algorithm~CITE, produce nothing more than a ``yes'', ``no'' or
``unknown'' response. Not only is this devoid of \emph{explanation}, but it also
hides the effects of any bugs; requiring the user to either trust the results,
or verify the implementation.

Both of these issues can be mitigated by having the system instead generate a
\emph{proof object}: a formal argument for \emph{why} a given statement follows
or does not follow. Once generated, a proof object's validity can be checked
without requiring any knowledge of how it was created, thus avoiding the need to
trust or verify the (potentially complicated) search and generation
procedures. The \emph{de Bruijn criterion}~CITE describes theorem provers which
produce proof objects that are trivial to check by independent \emph{proof
  checker} programs (which are themselves easily verified, due to their
simplicity); examples are Coq~CITE and Isabelle~CITE.

Proof objects are not a complete solution to understandability, since they can
still be quite inscrutable to human users. This often depends on how closely
the chosen formal system is able to encode the user's ideas: for example, the
formal proof of Kepler's Conjecture was performed using a system of Higher Order
Logic (HOL)~CITE whose proof objects (natural-deduction style derivations),
whilst tedious, are in principle understandable to a user experienced with both
the software and problem domain. The same cannot be said of the Boolean
Pythagorean Triples problem, a statement of Ramsey theory involving the
structure of the natural numbers. Rather than taking a high-level approach like
HOL, the authors of~CITE analysed sets $\{0..n\}$ for larger and larger $n$,
encoding these restrictions of the problem into the language of boolean
satisfiability (SAT), and finding that the problem is unsatisfiable for
$n= 7825$, and hence for the natural numbers as a whole. The proof object for
$n = 7825$ contains 200 terabytes (compressable to 68 gigabytes) of
propositional logic clauses, demonstrating the unsatisfiability of the encoded
problem instance. Not only is this far too much for any human to comprehend, but
the concepts used in the proof (boolean formulae) are several layers removed
from the actual problem statement (natural numbers, subsets and pythagorean
triples).

Whilst ``low level'' formalisms like SAT are less understandable or explanatory
for users, they are far more amenable to automation than more expressive logics.
Despite the proof for the Boolean Pythagorean Triples problem being many orders
of magnitude larger than that of the Kepler Conjecture, the latter is well
beyond the ability of today's automated theorem provers due to its encoding in
HOL. Instead, it took 22 collaborators 9 years just to formalise the proof
(Hales had previously produced a less formal proof, hundreds of pages long and
accompanied by unverified software).

I meant to send this earlier, in case it's useful. These are just some
things that came to mind regarding "explainable" AI and machine
learning. Lots of rambling, but hopefully some insight into current
machine-learning thinking (as I understand it, as somewhat of an
outsider!).

I think a key idea is "differentiable programming", which captures the
essence of how machine learning models like neural networks work, and
precisely what makes them different from traditional computer programs
(i.e. they're differentiable!), without ignoring the details ("black
box") or using flimsy analogies ("brain-like"):

https://fluxml.ai/2019/02/07/what-is-differentiable-programming.html

When run on an input, a differentiable program doesn't just give an
output, it also gives us the partial derivatives of that output
w.r.t. each of the program's parameters. This tells us exactly how to
adjust those parameters (via backpropagation) to reduce the output error
for that particular input. Doing this enough times for enough inputs
will hopefully find a general solution.

A neural network is an example of a differentiable program, one which is
just a big collection of identical expressions ("neurons") composed
together. This is incredibly general (they are universal function
approximators), but it doesn't have much "knowledge" or "structure"
baked in: the behaviour of this program depends almost completely on the
parameters, so everything it "learns" must come from input/output
examples.

The convolutional neural networks that have come to dominate image
processing are also differentiable programs, but they contain more
knowledge/structure. Instead of general-purpose 'neurons', these use
'convolutional layers' which are also expressions that get composed
together. The difference is that these expressions hard-code the 2D grid
structure that we expect to find in raster images. This makes them less
general than neural networks, since they assume the data fits this 2D
grid structure, but they also don't have to learn that structure from
examples. Hence they need fewer examples, but we can also be more
confident that the input/output relationships they learn involve 2D
patches of image content (like shapes or textures), rather than some
arbitrary numerical relationship between pixels.

Taking things much further, there are differentiable programs which do
ray tracing ( https://people.csail.mit.edu/tzumao/diffrt/ ). These can
be used to learn the position of 3D objects in an image, the camera
position, illumination, etc. Again, this is less general since it
assumes the input is an image of a 3D scene, but again it doesn't have
to learn this knowledge from examples. The learned parameters are easily
"explained", as they are the usual rendering parameters of camera
position, rotation, etc.

Another example is learning to recognise and generate hand-written
characters by using a differentiable program for drawing lines:
https://web.mit.edu/cocosci/Papers/Science-2015-Lake-1332-8.pdf

In this way, there are two tradeoffs being made:

 - By hard-coding our knowledge of the domain into a program, like using
   a ray-tracer for images of the 3D world, or pen-based drawing for
   images of hand-written characters, we don't need as much training
   data for the system to 'learn' what's going on. Unfortunately, we end
   up with less general systems, e.g. we might apply the same
   neural-network architecture to all sorts of problems (with enough
   data), but we can't apply the ray-tracer to e.g. audio data.

 - This same tradeoff also has implications for understandability and
   explainability: the parameters of a ray-tracer are directly
   understandable, based on how they're used in the program, e.g. "how
   bright is this light source" or "how big is this object". In
   contrast, the parameters of something like a neural network make
   little sense on their own, they just mean "how important is this
   output for that neuron?".

STUFF BELOW IS MORE TANGENTIAL

There are probably parallels to be drawn w.r.t. feature engineering
too. "Feature extraction" turns raw data into numbers for the learning
algorithm. This used to be done by hand, so we knew what the numbers
meant (e.g. average brightness, number of edges, etc. this is what
Katya's 'recurrent clustering' algorithm does for symbolic expressions).
With "deep nets", such features are figured out automatically, and we
don't know what they mean. Hence we gain generality (automatically
discovered features tend to out-perform hand-engineered ones), but we
lose understanding/explainability, and we become more dependent on
training data.

Things like "deep dream" can try to demonstrate what these features
seem to represent, but not much more. As far as I understand it, deep
dream works by picking a particular neuron/feature, providing an
arbitrary input image to the system, and iteratively tweaking that image
so that feature increases. The "explanation" of that feature is the
changes that were made to the image, e.g. if the changes look
eye-shaped, then we say that feature is for detecting eyes. This is, of
course, open to interpretation.

Interestingly, exactly the same technique is used for generating
adversarial inputs, where an input (usually to a classifier) is tweaked
until it gets classified incorrectly:

% https://en.wikipedia.org/wiki/Adversarial_machine_learning

Adversarial inputs show how limited this sort of "explanation" is, since
these changes might look like random noise:
https://blog.openai.com/adversarial-example-research/

Or might even just be a single pixel:
https://arxiv.org/abs/1710.08864

Another nice example of "explaining" a machine learning model by
demonstrating what its features represent is the variational
autoencoder.

Normal autoencoders 'compress' inputs down to a sequence of numbers
representing their important features. Unfortunately these are hard to
understand in isolation (sometimes called "holographic reduced
representations", all of the components depend on all of the others).

Variational autoencoders compress down to a multi-dimensional gaussian
distribution instead. By randomly sampling from this gaussian, the
representation is forced to be robust against small changes, and this
imposes a meaningful structure on the space; such that adjusting the
centre of the gaussian will alter the resulting data, e.g. by making a
person's hair darker; or changing their gender; or giving them glasses;
etc.

https://towardsdatascience.com/intuitively-understanding-variational-autoencoders-1bfe67eb5daf
(Especially the "Vector arithmetic" section)

Although this property can't actually be relied upon!
http://paulrubenstein.co.uk/variational-autoencoders-are-not-autoencoders/

A similar effect to the structured "latent space" of the VAE has been
found with word2vec:

% https://en.wikipedia.org/wiki/Word2vec#Preservation_of_semantic_and_syntactic_relationships

word2vec turns words into vectors of numbers, based on co-occurrences of
words in some training data. These vectors turn out to have some
"meaning" in the sense that vector arithmetic can make semantic changes.
For example if we do the sum:

  vector("king") - vector("man") + vector("woman")

The resulting vector is very close to vector("queen").

These sorts of machine learning system are unsupervised, so they don't
solve any particular problem. However, the representations they learn
can be useful for other task-specific systems.

\bibliographystyle{plain}
\bibliography{./Bibtex}

\end{document}