### 2019-02-28 UnderstandabilityPaper/machinelearning.tex
Finished draft of understandability

% STRUCTURE:

%  - Understandability in AI
%  - Theorem proving as example:
%   - Least informative is a yes/no/unknown answer
%   -

\documentclass{article}
\begin{document}
The problem of making machine-generated artefacts more understandable to human
users is quite clear in the field of \emph{automated theorem
  proving}~\cite{sutcliffe2001evaluating}: the task of deciding whether a given
formal statement follows from a given set of premeses.  The least informative
approaches, such algorithms based on the popular \emph{resolution}
rule~\cite{robinson1965machine}, produce nothing more than a ``yes'', ``no'' or
``unknown'' response. Not only is this devoid of \emph{explanation}, but it also
hides the effects of any bugs; requiring the user to either trust the results,
or verify the implementation.

Both of these issues can be mitigated by having the system instead generate a
\emph{proof object}: a formal argument for \emph{why} a given statement follows
or does not follow. Once generated, a proof object's validity can be checked
without requiring any knowledge of how it was created, thus avoiding the need to
trust or verify the (potentially complicated) search and generation
procedures. The \emph{de Bruijn criterion}~\cite{barendregt2005challenge}
describes theorem provers which produce proof objects that are trivial to check
by independent \emph{proof checker} programs (which are themselves easily
verified, due to their simplicity); examples are Coq~\cite{barras1997coq} and
Isabelle~\cite{nipkow2002isabelle}.

Proof objects are not a complete solution to understandability, since they can
still be quite inscrutable to human users. This often depends on how closely the
chosen formal system is able to encode the user's ideas: for example, the formal
proof of the Kepler Conjecture was performed using a system of Higher Order
Logic (HOL)~\cite{hales2015formal} whose proof objects (natural-deduction style
derivations), whilst tedious, are in principle understandable to a user
experienced with both the software and problem domain. The same cannot be said
of the Boolean Pythagorean Triples problem, a statement of Ramsey theory
involving the structure of the natural numbers. Rather than taking a high-level
approach like HOL, the authors of~\cite{heule2016solving} analysed sets
$\{0..n\}$ for larger and larger $n$, encoding these restricted versions of the
problem into the language of boolean satisfiability (SAT), and found that the
problem is unsatisfiable for $n= 7825$, and hence for the natural numbers as a
whole. In this case, the proof object demonstrates this unsatisfiability using
200 terabytes of propositional logic clauses (compressable to 68 gigabytes).
Not only is this far too much for any human to comprehend, but the concepts used
in the proof (boolean formulae) are several layers removed from the actual
problem statement (natural numbers, subsets and pythagorean triples).

Whilst ``low level'' formalisms like SAT are less understandable or explanatory
for users, they are far more amenable to automation than more expressive logics.
Despite the proof for the Boolean Pythagorean Triples problem being many orders
of magnitude larger than that of the Kepler Conjecture, the latter is well
beyond the ability of today's automated theorem provers due to its encoding in
HOL. Instead, it took 22 collaborators 9 years just to formalise the proof
(Hales had previously produced a less formal proof, hundreds of pages long and
accompanied by unverified software; a reminder that human-generated artefacts
are not necessarily understandable either!).

% A similar pattern has emerged when it comes to understanding

% I meant to send this earlier, in case it's useful. These are just some
% things that came to mind regarding "explainable" AI and machine
% learning. Lots of rambling, but hopefully some insight into current
% machine-learning thinking (as I understand it, as somewhat of an
% outsider!).

% I think a key idea is "differentiable programming", which captures the
% essence of how machine learning models like neural networks work, and
% precisely what makes them different from traditional computer programs
% (i.e. they're differentiable!), without ignoring the details ("black
% box") or using flimsy analogies ("brain-like"):

% https://fluxml.ai/2019/02/07/what-is-differentiable-programming.html

% When run on an input, a differentiable program doesn't just give an
% output, it also gives us the partial derivatives of that output
% w.r.t. each of the program's parameters. This tells us exactly how to
% adjust those parameters (via backpropagation) to reduce the output error
% for that particular input. Doing this enough times for enough inputs
% will hopefully find a general solution.

% A neural network is an example of a differentiable program, one which is
% just a big collection of identical expressions ("neurons") composed
% together. This is incredibly general (they are universal function
% approximators), but it doesn't have much "knowledge" or "structure"
% baked in: the behaviour of this program depends almost completely on the
% parameters, so everything it "learns" must come from input/output
% examples.

% The convolutional neural networks that have come to dominate image
% processing are also differentiable programs, but they contain more
% knowledge/structure. Instead of general-purpose 'neurons', these use
% 'convolutional layers' which are also expressions that get composed
% together. The difference is that these expressions hard-code the 2D grid
% structure that we expect to find in raster images. This makes them less
% general than neural networks, since they assume the data fits this 2D
% grid structure, but they also don't have to learn that structure from
% examples. Hence they need fewer examples, but we can also be more
% confident that the input/output relationships they learn involve 2D
% patches of image content (like shapes or textures), rather than some
% arbitrary numerical relationship between pixels.

% Taking things much further, there are differentiable programs which do
% ray tracing ( https://people.csail.mit.edu/tzumao/diffrt/ ). These can
% be used to learn the position of 3D objects in an image, the camera
% position, illumination, etc. Again, this is less general since it
% assumes the input is an image of a 3D scene, but again it doesn't have
% to learn this knowledge from examples. The learned parameters are easily
% "explained", as they are the usual rendering parameters of camera
% position, rotation, etc.

% Another example is learning to recognise and generate hand-written
% characters by using a differentiable program for drawing lines:
% https://web.mit.edu/cocosci/Papers/Science-2015-Lake-1332-8.pdf

% In this way, there are two tradeoffs being made:

%  - By hard-coding our knowledge of the domain into a program, like using
%    a ray-tracer for images of the 3D world, or pen-based drawing for
%    images of hand-written characters, we don't need as much training
%    data for the system to 'learn' what's going on. Unfortunately, we end
%    up with less general systems, e.g. we might apply the same
%    neural-network architecture to all sorts of problems (with enough
%    data), but we can't apply the ray-tracer to e.g. audio data.

%  - This same tradeoff also has implications for understandability and
%    explainability: the parameters of a ray-tracer are directly
%    understandable, based on how they're used in the program, e.g. "how
%    bright is this light source" or "how big is this object". In
%    contrast, the parameters of something like a neural network make
%    little sense on their own, they just mean "how important is this
%    output for that neuron?".

% STUFF BELOW IS MORE TANGENTIAL

% There are probably parallels to be drawn w.r.t. feature engineering
% too. "Feature extraction" turns raw data into numbers for the learning
% algorithm. This used to be done by hand, so we knew what the numbers
% meant (e.g. average brightness, number of edges, etc. this is what
% Katya's 'recurrent clustering' algorithm does for symbolic expressions).
% With "deep nets", such features are figured out automatically, and we
% don't know what they mean. Hence we gain generality (automatically
% discovered features tend to out-perform hand-engineered ones), but we
% lose understanding/explainability, and we become more dependent on
% training data.

% Things like "deep dream" can try to demonstrate what these features
% seem to represent, but not much more. As far as I understand it, deep
% dream works by picking a particular neuron/feature, providing an
% arbitrary input image to the system, and iteratively tweaking that image
% so that feature increases. The "explanation" of that feature is the
% changes that were made to the image, e.g. if the changes look
% eye-shaped, then we say that feature is for detecting eyes. This is, of
% course, open to interpretation.

% Interestingly, exactly the same technique is used for generating
% adversarial inputs, where an input (usually to a classifier) is tweaked
% until it gets classified incorrectly:

% % https://en.wikipedia.org/wiki/Adversarial_machine_learning

% Adversarial inputs show how limited this sort of "explanation" is, since
% these changes might look like random noise:
% https://blog.openai.com/adversarial-example-research/

% Or might even just be a single pixel:
% https://arxiv.org/abs/1710.08864

% Another nice example of "explaining" a machine learning model by
% demonstrating what its features represent is the variational
% autoencoder.

% Normal autoencoders 'compress' inputs down to a sequence of numbers
% representing their important features. Unfortunately these are hard to
% understand in isolation (sometimes called "holographic reduced
% representations", all of the components depend on all of the others).

% Variational autoencoders compress down to a multi-dimensional gaussian
% distribution instead. By randomly sampling from this gaussian, the
% representation is forced to be robust against small changes, and this
% imposes a meaningful structure on the space; such that adjusting the
% centre of the gaussian will alter the resulting data, e.g. by making a
% person's hair darker; or changing their gender; or giving them glasses;
% etc.

% https://towardsdatascience.com/intuitively-understanding-variational-autoencoders-1bfe67eb5daf
% (Especially the "Vector arithmetic" section)

% Although this property can't actually be relied upon!
% http://paulrubenstein.co.uk/variational-autoencoders-are-not-autoencoders/

% A similar effect to the structured "latent space" of the VAE has been
% found with word2vec:

% % https://en.wikipedia.org/wiki/Word2vec#Preservation_of_semantic_and_syntactic_relationships

% word2vec turns words into vectors of numbers, based on co-occurrences of
% words in some training data. These vectors turn out to have some
% "meaning" in the sense that vector arithmetic can make semantic changes.
% For example if we do the sum:

%   vector("king") - vector("man") + vector("woman")

% The resulting vector is very close to vector("queen").

% These sorts of machine learning system are unsupervised, so they don't
% solve any particular problem. However, the representations they learn
% can be useful for other task-specific systems.

\bibliographystyle{plain}
\bibliography{./bibtex}

\end{document}