### 2019-03-14 Bucketing/intro.tex
A bunch of bucketing metadata, and new intro

\section{Introduction}

\emph{Theory exploration} (TE) describes an open-ended search for
interesting properties (conjectured or proven) of a given set of formal
definitions, such as a mathematical theory, a scientific model or a software
library. It has been applied to generating ``background lemmas'' in the
interactive proof assistant \textsc{Isabelle/HOL}~TODO{CITE ISACOSY AND HIPSTER}
and for automated theorem proving~TODO{CITE PAPER WHICH SHOWS IMPROVED PROVING},
as well as for generating specifications for (pure functional) programming
libraries~TODO{CITE QUICKSPEC 1, 2 AND SPECULATE}.

Despite the clever search strategies employed by these tools, they all rely on
enumerating combinations of definitions in some way, which can cause infeasible
running times on larger inputs (in our experience, a few dozen definitions). To
avoid this, users of these systems must carefully select only small subsets of
their definitions to explore at a time.

We dub such selection the \emph{bucketing problem}, and analyse its effect on
the running time of theory exploration tools on a particular problem set, and on
the quality of the statements they are able to generate. An automated solution
to the bucketing problem would accept arbitrarily many definitions, and
efficiently select sub-sets which are small enough to reasonably explore whilst
still enabling interesting properties to be discovered.

Our contributions are:

\begin{enumerate}
  \item Defining the \emph{bucketing problem} for theory exploration.
  \item Statistical analysis of theory exploration performance on a large
    problem set, derived from existing theorem proving problem sets.
  \item Empirical analysis of the effect of bucketing on the quantity of
    desirable properties attainable by theory exploration.
  \item Experimental application of unsupervised machine learning to the
    bucketing problem.
  % \item A novel feature extraction method for transforming Haskell expressions
  %   into a form amenable to off-the-shelf learning algorithms.
  % \item An implementation of these feature extraction and theory exploration
  %   approaches.
  % \item A comparison of our methods with existing approaches, both for theory
  %   exploration in Haskell, and for machine learning in other languages.
\end{enumerate}

We describe theory exploration in more detail, as well as the \qspec{} tool we
use in our experiments, in \S~\ref{sec:theoryexploration}. We introduce the
bucketing problem in \S~\ref{sec:bucketing} and measure its effect on theory
exploration systems in \S~\ref{sec:implementation}. We attempt (unsuccessfully)
to apply machine learning to the bucketing problem in \S~\ref{sec:clustering}.
A variety of related work is surveyed in \S~\ref{sec:related} and we conclude in
\S~\ref{sec:conclusion} with potential directions for future research.