### 2019-04-09 Bucketing/sizes.tex
Use vector output

\begin{figure}
  \input{images/stepped.pgf}
\end{figure}

\begin{figure}
  \input{images/timeouts.pgf}
\end{figure}