### 2019-04-10 Bucketing/sizes.tex
Added captions

\begin{figure}
  \input{images/stepped.pgf}
  \caption{Survival plot % TODO: NAME OF PLOT
    for running QuickSpec on inputs containing various numbers of definitions,
    sampled from TIP. Most samples finished quickly, leaving a heavy tail which
    didn't finish within the 300 second timeout.}
  \label{fig:survival}
\end{figure}

\begin{figure}
  \input{images/timeouts.pgf}
  \caption{Proportion of samples which timed out per size, with least-squares
    linear regression.}
  \label{fig:tailsize}
\end{figure}