### 2019-04-11 Bucketing/sizes.tex
Add problematic definitions

\begin{figure}
  \input{images/stepped.pgf}
  \caption{Survival plot % TODO{2019-04-15}: NAME OF PLOT
    for running QuickSpec on inputs containing various numbers of definitions,
    sampled from TIP. Most samples finished quickly, leaving a heavy tail which
    didn't finish within the 300 second timeout.}
  \label{fig:survival}
\end{figure}

\begin{figure}
  \input{images/timeouts.pgf}
  \caption{Proportion of samples which timed out per size, with least-squares
    linear regression.}
  \label{fig:tailsize}
\end{figure}

\begin{figure}
  \input{images/proportions.pgf}
  \caption{}
  \label{fig:proportions}
\end{figure}

Five definitions appeared only in failing samples, these are named
\texttt{mult2}, \texttt{qexp}, \texttt{op}, \texttt{mul3acc} and \texttt{mul3};
their definitions appear in Figure~\ref{fig:faildefs}. All of these are
functions of natural numbers (\texttt{Nat}). The \texttt{mul3} and
\texttt{mul3acc} functions are rather pathological implementations of
multiplication with an accumulator parameter, involving many (non-tail)
recursive calls which can cause space leaks (or stack overflow, in languages
with eager evaluation). This is likely the reason \textsc{QuickSpec} performs
poorly with these definitions in the input.

The \texttt{op}, \texttt{mult2} and \texttt{qexp} functions are less easily
dismissed since they only recurse in tail position; the latter are in fact
standard definitions of multiplication and exponentiation, respectively. One
potential problem with these definitions, when translated into Haskell for
\textsc{QuickSpec}, is that their accumulating parameter will be lazy by
default, and hence cause space leaks.

\begin{figure}
  \texttt{
    (define-fun-rec mult2 ((x Nat) (y Nat) (z Nat)) Nat
      (match x
        (case  Z z)
        (case (S x2)
          (mult2 x2 y (plus y z)))))
  }

  \texttt{
    (define-fun-rec qexp ((x Nat) (y Nat) (z Nat)) Nat
      (match y
        (case  Z z)
        (case (S n)
          (qexp x n (mult x z)))))
  }

  \texttt{
    (define-fun-rec op ((x Nat) (y Nat) (z Nat) (x2 Nat)) Nat
      (match x
        (case Z
          (match z
            (case  Z x2)
            (case (S x3)
              (op Z y x3 (S x2)))))
        (case (S x4)
          (match z
            (case Z
              (op x4 y y x2))
            (case (S c)
              (op x y c (S x2)))))))
  }

  \texttt{
    (define-fun-rec mul3acc ((x Nat) (y Nat) (z Nat)) Nat
      (match x
        (case  Z Z)                         ;; Base case for 0 * y * z
        (case (S x2)
          (match y
            (case  Z Z)                     ;; Base case for x * 0 * z
            (case (S x3)
              (match z
                (case  Z Z)                 ;; Base case for x * y * 0
                (case (S x4)
                  (match x2
                    (case Z
                      (match x3
                        (case Z
                          (match x4
                            (case Z (S Z))  ;; Base case for 1 * 1 * 1
                            (case (S x5)
                              (S (add3acc (mul3acc Z Z x4)
                                          (add3acc (mul3acc (S Z) Z x4)
                                                   (mul3acc Z (S Z) x4)
                                                   (mul3acc Z Z (S Z)))
                                          (add3acc Z Z x4))))))
                        (case (S x6)
                          (S (add3acc (mul3acc Z x3 x4)
                                      (add3acc (mul3acc (S Z) x3 x4)
                                               (mul3acc Z (S Z) x4)
                                               (mul3acc Z x3 (S Z)))
                                      (add3acc Z x3 x4))))))
                    (case (S x7)
                      (S (add3acc (mul3acc x2 x3 x4)
                                  (add3acc (mul3acc (S Z) x3 x4)
                                           (mul3acc x2 (S Z) x4)
                                           (mul3acc x2 x3 (S Z)))
                                  (add3acc x2 x3 x4))))))))))))
  }

  \texttt{
    (define-fun-rec mul3 ((x Nat) (y Nat) (z Nat)) Nat
      (match x
        (case Z Z)                    ;; Base case for 0 * y * z
        (case (S x2)
          (match y
            (case Z Z)                ;; Base case for x * 0 * z
            (case (S x3)
              (match z
                (case Z Z)            ;; Base case for x * y * 0
                (case (S x4)
                  (match x2
                    (case Z
                      (match x3
                        (case Z
                          (match x4
                            (case Z
                              (S Z))  ;; Base case for 1 * 1 * 1
                            (case (S x5)
                              (S (add3 (mul3 Z Z x4)
                                       (add3 (mul3 (S Z) Z x4)
                                             (mul3 Z (S Z) x4)
                                             (mul3 Z Z (S Z)))
                                       (add3 Z Z x4))))))
                        (case (S x6)
                          (S (add3 (mul3 Z x3 x4)
                                   (add3 (mul3 (S Z) x3 x4)
                                         (mul3 Z (S Z) x4)
                                         (mul3 Z x3 (S Z)))
                                   (add3 Z x3 x4))))))
                    (case (S x7)
                      (S (add3 (mul3 x2 x3 x4)
                               (add3 (mul3 (S Z) x3 x4)
                                     (mul3 x2 (S Z) x4)
                                     (mul3 x2 x3 (S Z)))
                               (add3 x2 x3 x4))))))))))))
  }
  \label{fig:faildefs}
\end{figure}