### 2019-04-26 Bucketing/implementation.tex
More fleshing out

\section{Implementation}
\label{sec:implementation}

We provide an implementation of our recurrent clustering algorithm in a tool
called \textsc{ML4HS}. We obtain Core ASTs from Haskell packages using a plugin
for the GHC compiler, which emits a serialised version of each Core definition
as it is being compiled. This approach is more robust than, for example, parsing
source files, since it avoids the complications of preprocessors and language
extensions altering the syntax.

A post-processing stage determines which Core definitions can be used by
\qspec{}, based on their type, visibility (whether they are encapsulated inside
their module or visible to importers), etc. For each definition, we simply use
the GHCi interpreter to check if calling \qspec{} with that argument is
well-typed. This information, along with type signatures, arity, etc. are stored
alongside the Core definitions in JSON format.

The definitions are then sorted topologically, based on the non-local
identifiers appearing in their ASTs, and feature vectors are constructed using a
Haskell implementation of the approach described in \S
\ref{sec:recurrentclustering}. Since the features associated with each
identifier may vary between iterations (as the clusters change), we leave the
raw identifiers in the vector so their features can be extracted in the correct
context.

We implement Algorithm \ref{alg:recurrent} as a simple shell script, which
invokes Weka for clustering and associates each Core definition with a cluster
number. These numbers are used to finish the deferred feature extraction of
identifiers, the resulting feature vectors are clustered, and the process is
repeated until all SCCs have been processed.

Once the recurrent clustering is complete, we generate a string of Haskell code
for each of the final clusters, which will invoke \qspec{} with a suitable
signature. This involves:

\begin{itemize}
\item{Monomorphising}: If a value has polymorphic type, e.g. \hs{equal :: forall
    t. t -> t -> Bool}, we must choose a concrete representation to use (in this
  case for \hs{t}), in order to know which random generator to use. We take the
  approach used in \qcheck{} by attempting to instantiate all type variables to
  \hs{Integer}. Any values where this is invalid, such as those with
  incompatible class constraints (e.g. \hs{forall t. IsString t => t}, where
  \hs{Integer} does not implement \hs{IsString}) will not be included in the
  signature (this is checked at the AST post-processing stage).

\item{Qualifying}: All names are \emph{qualified} (prefixed by their module's
  name), to avoid most ambiguity. There is still the possibility that multiple
  packages will declare modules of the same name, although this is rare as it
  causes problems for any Haskell programmer trying to use those modules. In
  such cases the exploration process simply aborts.

\item{Appending variables}: Once a \qspec{} theory has been defined containing
  all of the given terms, we inspect the types it references and append three
  variables for each to the theory (enough to discover laws such as
  associativity, but not too many to overflow the limit of \qspec{}'s exhaustive
  search).

\item{Sandboxing}: One difficulty with Haskell's packaging infrastructure is
  that all required packages and modules must be provided up-front, usually by
  specification in a Cabal file. Since \textsc{MLSpec} builds signatures
  \emph{dynamically}, depending on the cluster information it is given, we do
  not know what packages it may need. To work around this problem, we evaluate
  these strings of Haskell using a custom library called \texttt{nix-eval}. This
  uses the Nix package manager to obtain all of the required packages and make
  them available to GHC.

\end{itemize}

The equations resulting from evaluating these strings are collected and
outputted as JSON values, to ease further processing (e.g. displaying in some
form, sending to a theorem prover, etc.).