### 2019-04-26 Bucketing/related.tex

\section{Related Work}
\label{sec:related}

\subsection{Theory Exploration}

We briefly described theory exploration in \S \ref{sec:theoryexploration}, as
the task of discovering \emph{new} theorems in a software or proof library,
rather than proving/disproving user-provided statements. This general idea dates
back at least to the 1970s, most famously to work by Lenat
\cite{lenat1977automated}, though the theory exploration tools used in
functional programming trace their roots back to the interactive
\textsc{Theorema} \cite{buchberger2000theory} system of Buchberger.

As well as \qspec{} and \hspec{} in Haskell, automated theory exploration has
been applied to Isabelle \cite{Montano-Rivas.McCasland.Dixon.ea:2012}
\cite{johansson2009isacosy} \cite{Hipster}. Since \qspec{} is used for
conjecture generation in many of these systems, they may also benefit from our
machine learning approach.

\subsection{Relevance Filtering}
\label{sec:relevance}

The idea of \emph{relevance filtering} (or, \emph{premise selection}) is to
reduce the search space of a theorem proving task by removing redundant or
irrelevant information (axioms, definitions, lemmas, etc.) from the input. Our
restriction of theory exploration to intelligently-selected clusters of symbols,
rather than whole libraries at a time, is essentially the same idea applied in a
novel setting.

The technique is used in Isabelle's Sledgehammer tool, during its translation of
Isabelle/HOL theories to statements in first order logic: rather than
translating the entire theory, only a sub-set of relevant clauses are
included. This reduces the size of the problem and speeds up the proof search,
but it creates the new problem of determining when a clause is relevant: how do
we know what will be required, before we have the proof?

The original approach uses the proportion of symbols which a clause has in
common with the problem statement \cite{meng2009lightweight} (plus various
heuristics), although various alternatives have been proposed in the literature
\cite{kuhlwein2013mash} \cite{kuhlwein2012overview} \cite{alama2014premise}.

\subsection{Recurrent Clustering}
\label{sec:clusteringexpressions}

Our recurrent clustering algorithm is similar to that of ML4PG, as our
transformation maps the elements of a syntax tree to distinct cells in a
matrix. In contrast, the matrices produced by ACL2(ml) \emph{summarise} the tree
elements: providing, for each level of the tree, the number of variables,
nullary symbols, unary symbols, etc.

Unlike ML4PG, our algorithm can handle mutually-recursive definitions, but it
also ignores types. This is because obtaining the type of each component of a
Core definition is more difficult than in ML4PG (which queries the current
\textsc{Proof General} state), and is left for future work.

The way we \emph{use} our clusters to inform theory exploration is actually more
similar to that of ACL2(ml) than ML4PG. ML4PG can either present clusters to the
user for inspection, or produce automata for recreating proofs. In ACL2(ml), the
clusters are used to restrict the search space of a proof search, much like we
restrict the scope of theory exploration.

\subsection{Feature Extraction}

One major difficulty when applying statistical machine learning algorithms to
\emph{languages}, such as Haskell, is the appearance of recursive
structures. This can lead to nested expressions of arbitrary depth, which are
difficult to compare in numerical ways. One common approach to this problem is
to represent such structures as \emph{sequences}. \emph{Recurrent neural
  networks} (RNNs) are a popular choice for processing sequences, especially
when combined with mechanisms such as \emph{long short-term memory} (LSTM) for
preserving information across long sequences \cite{hochreiter1997long}. Such
systems have been used, for example, to parse and execute computer programs
\cite{zaremba2014learning}. However, learning to parse sequences seems
inefficient considering that we already have correctly-parsed ASTs.

Whilst neural networks have been applied directly to recursive structures
\cite{goller1996learning}, including using LSTM \cite{zhu2015long}, a more
popular approach is to use \emph{kernel methods}
\cite{bakir2007predicting}. These are promising as a more principled alternative
to our current hand-crafted translation of ASTs to vectors.