### 2019-04-26 Bucketing/sizes.tex
More details of toxic definitions

\begin{figure}
  \scalebox{0.45}{\input{images/steppedall.pgf}}
  \scalebox{0.45}{\input{images/steppednontoxic.pgf}}
  \caption{Kaplan-Meier survival plot for running QuickSpec on inputs
    containing various numbers of definitions, sampled from TIP. The x axis
    denotes running time, which we cut short after 300 seconds. The height of
    each line shows the proportion of \qspec{} runs which were still going at
    that time (lower is better). First plot is for all TIP definitions, second
    plot removes runs given ``toxic'' definitions.}
  \label{fig:survival}
\end{figure}

Figure~\ref{fig:survival} shows a Kaplan-Meier survival plot of \qspec{} running
times, when given inputs containing different numbers of definitions. Many runs
finish quickly, with the remainder occupying a ``long tail'' which we cut off
after 300 seconds\footnote{Chosen based on preliminary experiments, which showed
  little difference in survival between 300 seconds and 1 hour}.

The number of definitions in the input is linearly correlated with the amount of
timeouts, except for the very smallest inputs (which are more constrained by the
sampling procedure). One explanation for this is ``toxic'' definitions, whose
presence in the input always leads to the exploration failing (with larger
inputs being more likely to have sampled a toxic definition).

\iffalse
\begin{figure}
  \scalebox{0.45}{\input{images/timeoutsall.pgf}}
  \scalebox{0.45}{\input{images/timeoutsnontoxic.pgf}}
  \caption{Proportion of samples which timed out per size, with least-squares
    linear regression. First plot is for all TIP definitions, second removes
    runs given ``toxic'' definitions.}
  \label{fig:tailsize}
\end{figure}
\fi

\begin{figure}
  \scalebox{0.45}{\input{images/proportionsall.pgf}}
  \scalebox{0.45}{\input{images/proportionsnontoxic.pgf}}
  \label{fig:proportions}
  \caption{Definitions, ordered by the ratio of successes to failures of the
    runs they appeared in. First graph contains all TIP definitions, showing
    ``toxic'' definitions which always failed. Second graph only contains runs
    without any toxic definitions.}
\end{figure}

\begin{figure}
  \begin{minted}{scheme}
    (define-fun-rec mult2 ((x Nat) (y Nat) (z Nat)) Nat
      (match x
        (case Z      z)
        (case (S x2) (mult2 x2 y (plus y z)))))
  \end{minted}

  \begin{minted}{scheme}
    (define-fun-rec qexp ((x Nat) (y Nat) (z Nat)) Nat
      (match y
        (case Z     z)
        (case (S n) (qexp x n (mult x z)))))
  \end{minted}

  \begin{minted}{scheme}
    (define-fun-rec op ((x Nat) (y Nat) (z Nat) (x2 Nat)) Nat
      (match x
        (case Z
          (match z
            (case Z      x2)
            (case (S x3) (op Z  y x3 (S x2)))))
        (case (S x4)
          (match z
            (case Z      (op x4 y y  x2))
            (case (S c ) (op x  y c  (S x2)))))))
  \end{minted}

  \iffalse
  \begin{minted}{scheme}
    (define-fun-rec mul3acc ((x Nat) (y Nat) (z Nat)) Nat
      (match x
        (case Z Z)                          ;; Base case for 0 * y * z
        (case (S x2)
          (match y
            (case Z Z)                      ;; Base case for x * 0 * z
            (case (S x3)
              (match z
                (case Z Z)                  ;; Base case for x * y * 0
                (case (S x4)
                  (match x2
                    (case Z
                      (match x3
                        (case Z
                          (match x4
                            (case Z (S Z))  ;; Base case for 1 * 1 * 1
                            (case (S x5)
                              (S (add3acc (mul3acc Z Z x4)
                                          (add3acc (mul3acc (S Z) Z x4)
                                                   (mul3acc Z (S Z) x4)
                                                   (mul3acc Z Z (S Z)))
                                          (add3acc Z Z x4))))))
                        (case (S x6)
                          (S (add3acc (mul3acc Z x3 x4)
                                      (add3acc (mul3acc (S Z) x3 x4)
                                               (mul3acc Z (S Z) x4)
                                               (mul3acc Z x3 (S Z)))
                                      (add3acc Z x3 x4))))))
                    (case (S x7)
                      (S (add3acc (mul3acc x2 x3 x4)
                                  (add3acc (mul3acc (S Z) x3 x4)
                                           (mul3acc x2 (S Z) x4)
                                           (mul3acc x2 x3 (S Z)))
                                  (add3acc x2 x3 x4))))))))))))
  \end{minted}
  \fi

  \iffalse
  \begin{minted}{scheme}
    (define-fun-rec mul3 ((x Nat) (y Nat) (z Nat)) Nat
      (match x
        (case Z Z)                          ;; Base case for 0 * y * z
        (case (S x2)
          (match y
            (case Z Z)                      ;; Base case for x * 0 * z
            (case (S x3)
              (match z
                (case Z Z)                  ;; Base case for x * y * 0
                (case (S x4)
                  (match x2
                    (case Z
                      (match x3
                        (case Z
                          (match x4
                            (case Z (S Z))  ;; Base case for 1 * 1 * 1
                            (case (S x5)
                              (S (add3 (mul3 Z Z x4)
                                       (add3 (mul3 (S Z) Z x4)
                                             (mul3 Z (S Z) x4)
                                             (mul3 Z Z (S Z)))
                                       (add3 Z Z x4))))))
                        (case (S x6)
                          (S (add3 (mul3 Z x3 x4)
                                   (add3 (mul3 (S Z) x3 x4)
                                         (mul3 Z (S Z) x4)
                                         (mul3 Z x3 (S Z)))
                                   (add3 Z x3 x4))))))
                    (case (S x7)
                      (S (add3 (mul3 x2 x3 x4)
                               (add3 (mul3 (S Z) x3 x4)
                                     (mul3 x2 (S Z) x4)
                                     (mul3 x2 x3 (S Z)))
                               (add3 x2 x3 x4))))))))))))
  \end{minted}
  \fi
  \caption{``Toxic'' definitions, which consistently cause \qspec{} to fail. Two
    other definitions (\texttt{mul3} and \texttt{mul3acc}) are ommitted due to
    their verbosity.}
  \label{fig:faildefs}
\end{figure}

This does appear to be the case, as Figure~\ref{fig:proportions} shows that five
definitions appeared \emph{only} in failing inputs; these are named
\texttt{mul3}, \texttt{mul3acc}, \texttt{mult2}, \texttt{op} and \texttt{qexp};
their definitions appear in Figure~\ref{fig:faildefs}. All of these are
functions of Peano-encoded natural numbers (\texttt{Nat}), and they cause
exploration to time out by either exhausting the RAM with too many expressions
to explore, % TODO{2019-04-27} CAN QUICKSPEC2 AVOID THIS?
or by generating such deeply-nested outputs that comparing them takes too long.
% TODO{2019-04-27} CAN WE PUT TIME AND MEMORY LIMITS ON EACH ATTEMPTED EVALUATION?

The \texttt{mul3} and \texttt{mul3acc} definitions are rather pathological
implementations of multiplication with an accumulator parameter, with many
(non-tail) recursive calls. The \texttt{op} function appears in files named
\texttt{weird\_nat\_op}, which assert its commutativity and associativity.
Finally, the \texttt{mult2} and \texttt{qexp} functions are standard
tail-recursive definitions of multiplication and exponentiation, respectively.
All of these functions have an extra ``accumulator'' argument, which increases
the number of possible expressions to explore compared to those without.

Exploring each of these functions on its own does not require much memory, since
Haskell generates the output lazily. However, comparing such large numbers for
equality takes a lot of CPU time as the \texttt{S} constructors are successively
unwrapped from each side, and this is why the timeout is reached. We confirmed
this hypothesis by exploring with a custom data generator which only generates
the values \texttt{Z}, \texttt{S Z} and \texttt{S (S Z)} (0, 1 and 2); this
caused the exploration to finish quickly. Other interventions, like making the
accumulator arguments strict (to prevent space leaks), did not prevent timeouts.

To assess the impact of these problematic definitions, we removed any samples
containing them and repeated our analysis.