### 2019-04-29 PhDThesis/litreview_machinelearning.tex

\subsection{Machine Learning Applications in Mathematics, Logic and Programming
  Language Theory}

Machine learning techniques, used for finding statistical patterns and numerical
correlations in arbitrary data, are characterised by their statistical
pattern-finding

% TODO: AutoTuning

% DeepTune analyses raw program text to predict the best options for an optimising
% compiler. FIXME

% TODO Fuzzing AFL DeepSmith