### 2019-04-29 PhDThesis/litreview.tex
Disable some bits of litreview for now

\section{Literature Review}

\input{litreview_automatedreasoning.tex}
\input{litreview_staticanalysis.tex}
\input{litreview_testing.tex}
\input{litreview_premiseselection.tex}
% TODO
%\input{litreview_machinelearning.tex}
%\input{litreview_interestingness.tex}