### 2019-05-16 PhDThesis/appendix_nixeval.tex
Appendix changes

\section{\texttt{nix-eval}}
\label{sec:nix-eval}

\begin{sloppypar}
  Haskell's support for runtime evaluation of code (\'a la Lisp's
  \eval{}~\cite[\S~1.6]{edwards1962lisp}) is limited. The most popular mechanism
  is the \hint{} package\footnote{https://hackage.haskell.org/package/hint},
  which is built on top of GHC's API and provides functions like
  \hs{eval :: MonadInterpreter m => String -> m String}. The
  \hs{MonadInterpreter} constraint provides mechanisms for calling into GHC and
  managing its ``session'', which includes details like compiler flags, a list
  of imported modules and paths to compiled packages.

  This function doesn't return arbitrary types of value; only \hs{String}s. This
  is due to the given expression being passed to the \hs{show} function (e.g.
  \hs{eval "not True"} will produce \hs{"False" :: String}, not
  \hs{False :: Bool}).\footnote{\hs{show :: Show a => a -> String} is a method
    of the \hs{Show} type class, and is the standard way to equip Haskell types
    with a printable representation (akin to Java's \texttt{toString} method).
    \hs{show} has no associated laws, but convention is for \hs{show} to be the
    inverse of the \hs{read :: Read a => String -> a} method of the \hs{Read}
    type class (where possible).} Serialising results in this way simplifies the
  interface between evaluation contexts (in particular, type checking and
  inference doesn't need to cross the interface); at the cost of having to
  explicitly deserialise results into the desired type as needed.

  One major problem with \hint{} is that the available packages
  (and GHC version) are hard-coded into the ``outer'' Haskell context. For
  example, a Haskell interpreter based on \hint{} would be restricted to only
  those packages and GHC features that are available during its
  compilation.\footnote{The package database could be overridden in principle,
    but compatibility issues and version conflicts would make this unreliable
    and error-prone in practice.} For our research we want a single program for
  exploring \emph{arbitrary} Haskell code, regardless of what package it comes
  from, which makes approaches like those of \hint{} undesirable.

  To solve this problem we created the \nixeval{}
  package\footnote{https://hackage.haskell.org/package/nix-eval}. This provides
  a function \hs{eval :: Expr -> IO (Maybe String)}, which produces \hs{String}s
  for the same reasons as \hint{}, but differs in its handling of expressions
  and context. Rather than operating on a raw \hs{String}, \nixeval{} defines
  the \hs{Expr} type, containing several fields to keep track of the context in
  which the expression is to be evaluated:
\end{sloppypar}

\begin{itemize}
\item \hs{eExpr :: String} is the raw code itself.
\item \hs{ePkgs :: [Pkg]} lists the packages which need to be installed.
\item \hs{eMods :: [Mod]} lists the modules which need to be imported.
\item \hs{eFlags :: [Flag]} lists flags which need passing to the compiler (e.g.
  to enable certain language features).
\item \hs{ePreamble :: [String]} lists code that needs to be defined before
  evaluating (for example, custom function definitions).
\end{itemize}

The types \hs{Pkg}, \hs{Mod} and \hs{Flag} are just wrappers containing a
\hs{String}; they are kept distinct to allow the type checker to catch
mistakes.\footnote{Avoiding such distinctions (i.e. using types to track
  representations but not semantics) is known as ``stringly-typed
  programming''~\cite{c2stringlytyped}. In the case of binary choices this
  practice is known as ``boolean blindness''~\cite[\S~26.2]{Harper:2009}.} These
wrapper types, as well as \hs{Expr} itself, implement the \hs{IsString} type
class, which allows GHC's \texttt{OverloadedStrings} language extension to parse
them from quoted literals, for example:

\begin{haskell}
textToUpper :: Expr
textToUpper = withPkgs       ["text"     ]  -- This is a  Pkg
                (withMods    ["Data.Text"]  -- This is a  Mod
                  (withFlags ["-O2"      ]  -- This is a  Flag
                    "toUpper"))             -- This is an Expr
\end{haskell}

The main advantage over \hint{} is the reification of packages, which can be
computed in arbitrary ways (including from user input). When an \hs{Expr} is
passed to our \hs{eval} function, a new \texttt{runhaskell} process is launched
with the given \hs{eFlags} (hence the presence of \hs{IO}). The elements of
\hs{eMods} are prefixed with \hs{"import "} and fed into its standard input
stream, followed by the contents of \hs{ePreamble}. The \hs{eExpr} code is
wrapped in parentheses and prefixed with \hs{"main = Prelude.putStr"} which acts
as the entry point of the generated code (writing the expression, which must be
a \hs{String}, to standard output).\footnote{The code generation for \hs{eExpr}
  is actually user-configurable, in case the default is too simplistic. Unlike
  \hint{} we never force calls to the \hs{show} method, since normal function
  calls are easier to specify and override than type class methods.} If the
subprocess exits successfully, its standard output is returned (wrapped in
\hs{Just}); otherwise it is written to the standard error stream and a
\hs{Nothing} value is returned (the subprocess inherits the parent's standard
error stream).

The \texttt{runhaskell} command can be invoked in two ways: if
\texttt{runhaskell} is available via the \texttt{PATH} environment variable, and
the contents of \hs{ePkgs} are all installed in its package database, then we
invoke that command directly. If not, we use the \texttt{nix-shell} command of
the Nix package manager to invoke \texttt{runhaskell} in a
sandbox~\cite{dolstra2004nix}. We use the Nix language to request a sandbox
containing GHC and all of the elements of \hs{ePkgs}. We also allow the default
Nix packages to be overridden via an environment variable, which allows the user
to define their own packages, select particular versions, or perform any other
arbitrary setup (since the Nix language is Turing-complete).

Nix packages are immutable and (mostly) reproducible, which allows sandboxes and
their components to be re-used, and even shared across machines. However, any
package which isn't cached in this way must be built from source, which can
cause \hs{eval} to take an arbitrarily long time. Even a cached sandbox can take
a few seconds to initialise (largely due to I/O latency as Nix computes and
inspects the sandbox components). To avoid too much overhead, it is advisable to
invoke \hs{eval} as few times as possible (ideally only once). If we need to
dynamically evaluate multiple parts of a program, it might be preferable to move
them all into one big \hs{Expr} value.

\begin{figure}
  \begin{haskell}
-- Apply x to n undefined arguments; result is undefined with x's output type
addCalls :: Int -> Expr -> Expr
addCalls 0 x = x
addCalls n x = addCalls (n-1) (x $$ "undefined")

-- Identity function, monomorphised to the output type of "f" (of arity a)
converter :: Expr
converter = ("flip" $$ "asTypeOf") $$ addCalls a "f"

-- Composes the given list of functions using the (.) operator
compose :: [Expr] -> Expr
compose [x]    = x
compose (x:xs) = ("(.)" $$ x) $$ compose xs

-- Forces a value's type, serialises it then hashes it
hasher :: Expr
hasher = compose [withPkgs ["murmur-hash"]
                    (qualified "Data.Digest.Murmur32" "asWord32"),
                  withPkgs ["murmur-hash"]
                    (qualified "Data.Digest.Murmur32" "hash32"),
                  withPkgs ["cereal"]
                    (qualified "Data.Serialize" "runPut"),
                  withPkgs ["cereal"]
                    (qualified "Data.Serialize" "put"),
                  converter]
  \end{haskell}
  \caption{Example usage of \nixeval{} functions, taken from \textsc{ML4HS}.
    These are used to dynamically generate \qspec{} signatures of user-provided
    functions.}
  \label{fig:nix-eval}
\end{figure}

\begin{sloppypar}
  Using \hs{String} functions to manipulate such complicated values soon becomes
  unwieldy, so a suite of combinators (such as the \hs{withFoo} functions above)
  are provided for composing \hs{Expr} values in useful ways; examples are shown
  in Figure~\ref{fig:nix-eval}. The \hs{qualified} function is particularly
  useful, for qualifying names with their module (reducing ambiguity and making
  composition more robust) and appending to \hs{eMods}, i.e.
  \hs{qualified "Data.Text" "toUpper"} is equivalent to
  \hs{withMods ["Data.Text"] "Data.Text.toUpper"}. Function calls are also very
  common, which can be written as \hs{foo $$ bar} (analogous to Haskell's
  standard \hs{$} %$
  operator); this wraps both \hs{Expr}essions in parentheses
  and juxtaposes them, appending both contexts.
\end{sloppypar}

Despite being more structured than raw \hs{String}s, the the above API still
suffers from a form of ``stringly-typed programming''; notably, that every
encoded expression has the same type: \hs{Expr}. We can avoid this through the
use of \emph{phantom types}: overly-restrictive type annotations, which are
unrelated to the encoding but must still match up in order to pass the type
checker~\cite[\S~5.1]{leijen1999domain}. In particular we can represent typed
expressions like this:

\begin{haskell}
newtype TypedExpr t = TE Expr
\end{haskell}

Notice that the parameter \hs{t} doesn't appear on the right hand side: the
values of \hs{TypedExpr} are just those of \hs{Expr} (the wrapper, along with
the phantom type, is \emph{erased} after type-checking). The \hs{TE} constructor
is hence polymorphic in \hs{t}, so \hs{TE x :: TypedExpr T} is valid for
\emph{any} choice of \hs{x} and \hs{T} (similar to how \hs{Nil :: List T} is
valid for any choice of \hs{T}). In particular, this includes the type of the
expression \emph{represented by} \hs{x}, for example
\hs{TE "True" :: TypedExpr Bool} is a type-correct expression.

Haskell will infer the most general type of an expression if none is provided
(i.e. \hs{forall t. TypedExpr t}), but will honour more restricted annotations
if given. This allows the types represented by literal expressions to be
asserted, for example:

\begin{haskell}
nothing' :: TypedExpr (Maybe a)
nothing' = TE (qualified "Data.Maybe" "Nothing")

just' :: TypedExpr (a -> Maybe a)
just' = TE (qualified "Data.Maybe" "Just")
\end{haskell}

Annotations on the combinator library can propagate and check these phantom
types, most notably:

\begin{haskell}
-- Typed function application
($$$) :: TypedExpr (a -> b) -> TypedExpr a -> TypedExpr b
TE f $$$ TE x = TE (f $$ x)

-- Only expressions of type String can be evaluated
tEval :: TypedExpr String -> IO (Maybe String)
tEval (TE x) = eval x
\end{haskell}

Haskell will not check that these annotation match the encoded expressions, so
care must be taken to ensure that they correspond correctly. Unannotated values
will remain polymorphic and unify with anything, which reduces the verbosity and
correctness burden (e.g. for intermediate values) whilst still guaranteeing
consistency (unification will fail if multiple incompatible types are used). If
user-provided literals aren't required then the \hs{TE} constructor can be kept
private to prevent unsound assertions or coercions.

However, there is a more subtle problem with such \hs{TypedExpr}s: the whole
point of \nixeval{} is to manipulate and evaluate code whose dependencies are
unknown or unavailable at compile time. For example, we cannot assign a
meaningful type at compile time to an expression which has been generated from
user input. Even if we know the value of an expression at compile time, its type
may come from a dependency that our program doesn't have access to. In these
cases the \hs{TypedExpr} approach can still be useful for auxiliary code whose
types are known to us; those unknown or unavailable parts can either be left
polymorphic, unwrapped into raw \hs{Expr}s, or dummy types can be used instead.
For example, if we are manipulating code involving the \hs{Data.Text.Text} type,
but we don't have or want the \texttt{text} package in our dependencies, we can
declare an empty type like:

\begin{haskell}
data TEXT  -- No constructors given, making this type empty
\end{haskell}

We can use this in our annotations in place of \hs{Data.Text.Text}. The compiler
will type-check our code as desired, ensuring functions are called on
appropriate arguments, \hs{tEval} is called with a \hs{String}, etc.

In the course of writing \mlspec{} we experimented with both raw \hs{Expr} and
\hs{TypedExpr} values. Whilst we found the latter easier for constructing large
expressions, we ultimately found it even more productive to move as much
functionality as possible into a standalone Cabal package
(\texttt{mlspec-helper}), add that package to the Nix package set and have
\mlspec{} call it using simple \hs{Expr} values.