### 2019-06-02 PhDThesis/background_theoryexploration.tex

\section{Theory Exploration}
\label{sec:theoryexploration}

(Automated) Theory Exploration (TE) is the task of taking a \emph{signature}
of definitions in some formal system (for example a programming language) and
automatically generating a set of formal statements (properties) involving
those definitions. These may be conjectures or theorems (proven either by
sending conjectures to an automated theorem prover, or by having the generating
procedure proceed in logically sound steps). These statements must also be
``interesting'' in some way, to rule out unhelpful trivialities such as
iterating a pattern over and over (e.g. \hs{x + 0 = x}, \hs{(x + 0) + 0 = x},
\dots).

\iffalse
It is beyond the scope of this paper
to define what makes a mathematical statement ``interesting'', but this problem
has been tackled extensively in the literature~\cite{colton2000notion}.
\fi
% END TOTO

The choice of appropriate methods can differ according to the underlying logic,
etc. that is in use, so in this work we ground ourselves by applying TE to
discovering properties of pure functional software libraries, mostly in the
Haskell programming language. This also gives our results immediate utility to
software engineering. We ignore the issue of \emph{proving} such conjectures,
deferring that task to existing tools (such as \hipspec{}).

The method of conjecture generation is a key characteristic of any theory
exploration system, although all existing implementations rely on brute force
enumeration to some degree. We focus on \quickspec{}~\cite{QuickSpec}, which
conjectures equations about Haskell code. We have used version 1 of
\quickspec{} in our experiments, due to its availability, stability and tooling
integration; hence references to \quickspec{} are to version 1 unless stated
otherwise. At the time of writing there is a \quickspec{} version 2 available,
which has a much improved generation procedure which can be significantly faster
than its predecessor. Whilst \quickspec{} 2 has advanced the state of the art in
theory exploration, our preliminary experience has found that it still suffers
the scaling issues we identify in this work\footnote{We encountered memory
  exhaustion when trying to run some of the examples included with the
  \quickspec{} 2 source code; let alone auto-generated signatures.} Since
\quickspec{} 2 has less integration with the tooling we have used, we leave a
more thorough analysis of it (and related systems such as \speculate{}) for
future work, pending the necessary infrastructure changes that would require.

\begin{figure}
  \centering
  \begin{minted}{haskell}
    -- A datatype with two constructors (Note that -- introduces a comment)
    data Bool = True | False

    -- A recursive datatype (S requires a Nat as argument)
    data Nat = Z | S Nat

    -- A function turning a Bool into a Bool
    not :: Bool -> Bool
    not True  = False
    not False = True

    -- A pair of mutually recursive functions

    odd :: Nat -> Bool
    odd    Z  = False
    odd (S n) = even n

    even :: Nat -> Bool
    even    Z  = True
    even (S n) = odd n
  \end{minted}
  \caption{Haskell datatypes for booleans and natural numbers, followed by some
    simple function definitions (with type annotations).}
  \label{fig:haskellteexample}
\end{figure}

\quickspec{} (version 1) is written in the Haskell programming language, and
works with definitions which are also written in Haskell. A thorough description
of Haskell, including its suitability for theory exploration, can be found
in \S~\ref{sec:haskell}. For illustrative purposes, some simple Haskell
definitions are shown in Figure~\ref{fig:haskellteexample}. \quickspec{} uses
the following procedure to generate conjectures, which are both plausible (due
to testing on many examples) and potentially interesting (due to being mutually irreducible):

\begin{enumerate}
\item Given a signature $\Sigma$ of typed expressions and set of variables $V$,
  \quickspec{} generates a list $terms$ containing the expressions from
  $\Sigma$ (including functions), the variables from $V$ and type-correct
  function applications \hs{f x}, where \hs{f} and \hs{x} are elements of
  $terms$. To ensure the list is finite, function applications are only nested
  up to a specified depth (by default, 3).
\item The elements of $terms$ are grouped into equivalence classes, based on
  their type.
\item The equivalence of terms in each class is tested using \quickcheck{}:
  variables are instantiated to particular values, generated randomly, and the
  resulting closed expressions are evaluated and compared for equality.
\item If a class is found to have non-equal members, it is split up to separate
  those members.
\item The previous steps of variable instantiation and comparison are repeated
  until the classes stabilise (i.e. no differences have been observed for some
  specified number of repetitions).
\item For each class, one member is selected and equations are conjectured that
  it is equal to each of the other members.
\item A congruence closure algorithm is applied to these equations, to discard
  any which are implied by the others.
\end{enumerate}

Such conjectures can be used in several ways: they can be presented directly
to the user, sent to a more rigorous system like \hipspec{} or \hipster{} for
proving, or even serve as a background theory for an automated theorem
prover~\cite{claessen2013automating}.

As an example, we can consider a simple signature containing the expressions
from Figure~\ref{fig:haskellteexample}, and some suitable variables:

\begin{align*}
  \Sigma_{\hs{Nat}} &= \{\hs{Z},\ \hs{S},\ \hs{plus},\ \hs{mult},\ \hs{odd},
                       \ \hs{even}\} \\
  V_{\hs{Nat}}     &= \{\hs{a :: Nat},\ \hs{b :: Nat},\ \hs{c :: Nat}\}
\end{align*}

\quickspec{}'s enumeration of these terms will resemble the following:

\begin{align*}
  terms_{\hs{Nat}} = [& \hs{Z},\ \hs{S},\ \hs{plus},\ \hs{mult},\ \hs{odd},
                     \ \hs{even},\ \hs{a},\ \hs{b},\ \hs{c},\ \hs{S Z},
                     \ \hs{S a},\ \hs{S b}, \\
                     & \hs{S c},\ \hs{plus Z},\ \hs{plus a},\ \dots ]
\end{align*}

Notice that functions such as \hs{plus} and \hs{mult} are valid terms, despite
not being applied to any arguments. In addition, all Haskell functions are unary
(due to currying), which makes it valid to apply them one argument at a time as
we construct $terms_{\hs{Nat}}$.

\begin{figure}
  % To reproduce, run 'quickSpec nat' in haskell_example/src/QuickSpecExample.hs
  \begin{haskell}
                      plus a b = plus b a
                      plus a Z = a
             plus a (plus b c) = plus b (plus a c)
                      mult a b = mult b a
                      mult a Z = Z
             mult a (mult b c) = mult b (mult a c)
                  plus a (S b) = S (plus a b)
                  mult a (S b) = plus a (mult a b)
             mult a (plus b b) = mult b (plus a a)
                     odd (S a) = even a
                odd (plus a a) = odd Z
               odd (times a a) = odd a
                    even (S a) = odd a
               even (plus a a) = even Z
              even (times a a) = even a
    plus (mult a b) (mult a c) = mult a (plus b c)
  \end{haskell}
  \caption{Equations conjectured by \quickspec{} for the functions in Figure
    \ref{fig:haskellteexample}; after simplification.}
  \label{fig:qspecresult}
\end{figure}

The elements of $terms_{\hs{Nat}}$ will be grouped into five classes, one each
for \hs{Nat}, \hs{Nat -> Nat}, \hs{Nat -> Nat -> Nat}, \hs{Nat -> Bool} and
\hs{Bool}. As the variables \hs{a}, \hs{b} and \hs{c} are instantiated to
various randomly-generated numbers, the differences between these terms will
be discovered and the equivalence classes will be divided, until eventually
the equations in Figure \ref{fig:qspecresult} are conjectured.

Although complete, this enumeration approach is wasteful: many terms are
unlikely to appear in theorems, which requires careful choice by the user of
what to include in the signature. Here we know that addition and multiplication
are closely related, and hence obey many algebraic laws; arbitrary definitions
from a typical software library or proof development are unlikely to have be
related so strongly.

\iffalse
\quickspec{} (and \hipspec{}) is also compatible with Haskell's existing testing
infrastructure, such that an invocation of \texttt{cabal test} can run these
tools alongside more traditional QA tools like \quickcheck{}, \textsc{HUnit} and
\textsc{Criterion}.

In fact, there are similarities between the way a TE system like \quickspec{} can
generalise from checking \emph{particular} properties to \emph{inventing} new
ones, and the way counterexample finders like \quickcheck{} can generalise from
testing \emph{particular} expressions to \emph{inventing} expressions to
test. One of our aims is to understand the implications of this generalisation,
the lessons that each can learn from the other's approach to term generation,
and the consequences for testing and QA in general.
\fi

\iffalse
% TODO: Work this into the example

As an example, giving the definitions from Figure~\ref{fig:haskellteexample} to
\quickspec{}, along with suitable comparison functions, random data generators and
variables \hs{n :: Nat} and \hs{b :: Bool}, will give rise to an equivalence
class for each type \hs{Nat}, \hs{Bool}, \hs{Nat -> Nat}, \hs{Nat -> Bool} and
\hs{Bool -> Bool}. Testing will find unequal terms in some of these classes
(such as the \hs{Bool} class containing \hs{True}, \hs{False}, \hs{b},
\hs{not b}, \hs{odd n} and \hs{even n}, which are all mutually unequal), split
them up, and repeat until no more splits are found. Equations are then generated
between elements of non-singleton classes, such as $\{\text{\hs{odd n}},
\text{\hs{not (even n)}}, \text{\hs{not (not (odd n))}}, \ldots\}$, and reduced
to a minimal set (for example, discarding \hs{odd n == not (not (odd n))} as an
instance of the more general \hs{b == not (not b)}). The remaining equations
(including \hs{not (not b) == b} and \hs{not (odd n) == even n}, etc.) are
output, either for the user to peruse or for another system to process, like the
\hipspec{} automated theorem prover.

Note that we must be able to generate values of a type in order to include
variables of that type. Also, only equivalence classes whose element type can be
compared are split and conjectured about in this way; in particular, functions
(like \hs{odd} and \hs{even}) are first-class values, so they will be put in an
equivalence class, but they will not be subject to testing and comparison since
Haskell has no generic notion of function equality~\footnote{In
  fact \quickspec{} compares expressions using a total ordering rather than equality;
  this is even more restrictive than requiring a notion of equality, but reduces
  the number of required comparisons}. \quickspec{} can hence conjecture that two
functions are pointwise-equal (e.g. \hs{f x == g x}) but it cannot conjecture
that the functions themselves are equal (e.g. \hs{f == g}).
\fi

% TODO
%%%%% THE BELOW SEEMS TO BE INTERESTINGNESS?

%\begin{description}
%\item{Interestingness} Various alternative interestingness criteria have been
%  proposed, which we survey in \S \ref{sec:relatedwork}. Augmenting or replacing
%  the criteria may be useful, for example to distinguish useful relationships
%  from incidental coincidences; or to prevent surprising, insightful equations
%  from being discarded because they can be simplified.
%\end{description}

% Theory exploration is similar to \emph{experimental mathematics}

% - Relation to Science
%  - Testable/falsifiable hypotheses are like evaluable terms (or, more generally, conjectures which can be decided, using a reasonable amount of resources).
% - Relation to AI tasks: exploring surroundings, etc.

% - Statistics is another area that's less straightforward than normal numerical computing, since there is subjectivity and judgement involved in the answering of questions.

% Theory formation: Alison? Others.
% Theory exploration: Buchberger, Moa in Isabelle, Koen in Haskell. Others?
% Theorem proving: Well-trodden: first-order ATP, higher-order ITP, functional programming
% Communication: Latex, Wikis, APIs, communicating with aliens

% \section{Exploration in Theorem Proving}
% \label{sec:examples}

% Before exploring abstract definitions of interestingness, we can first consider
% some scenarios which arise during formal proof where we are forced to generate
% conjectures. An analysis of these situations, and the subsequent theorems they
% produce, will contribute towards an empirical justification for what is
% interesting (at least from a utilitarian point of view) and inform our later
% exploration of the literature.

% \subsection{Generalisation}

% \providecommand{\coq}[1]{\lstinline[language=ML]|#1|}

% When we \emph{generalise} a statement $S$, we obtain a new statement $S'$ of
% which $S$ is a special case. Although it seems counterintuitive, a generalised
% statement can sometimes be \emph{easier} to prove than the original. This arises
% often in inductive proofs, since the specific obligations which arise in the
% proof may be incompatible with the available inductive hypotheses.

% However, we cannot blindly generalise \emph{all} obligations we encounter, since
% \emph{over-generalising} results in obligations which are so strong that they
% are unprovable, or even false. We must therefore rely on heuristics to guide the
% generation of generalised conjectures, and hence perform a kind of exploration.

% An informative example is given by Boyer and Moore of the associativity of
% multiplication in ACL2 \cite{boyer1983proof}:

% $$(x * y) * z = x * (y * z)$$

% During the course of the proof, the following obligation arises:

% \begin{equation}
%   \tag{conc3}
%   (y + (x * y)) * z = (y * z) + ((x * y) * z)
%   \label{eq:conc3}
% \end{equation}

% ACL2 automatically generalises \eqref{eq:conc3} by replacing the repeated
% sub-term $x * y$ with a fresh variable $w$:

% \begin{equation}
%   \tag{conc4}
%   (y + w) * z = (y * z) + (w * z)
%   \label{eq:conc4}
% \end{equation}

% This generalised form is clearly the distributivity law for multiplication and
% addition, which can be proved separately to the original goal of
% associativity. It would not be controversial to claim that this distributivity
% law is interesting in its own right (relative to associativity, at least), in
% addition to its usefulness in making this proof go through.

% % TODO: Describe the ACL2 heuristics

% Generalisation also occurs frequently when reasoning about \emph{tail-recursive}
% definitions \cite{kapur2003automatic}. \footnote{A tail-recursive function can
%   be executed in constant space using a loop, whereas recursion in non-tail
%   positions may require a growing number of stack frames or nested closures. See
%   \S \ref{sec:auxiliarylemmas} for example definitions of each type.}

% \subsection{Analogy}

% One way to characterise the interestingness of a statement is by \emph{analogy}
% to existing interesting statements. By finding lemmas analogous to those of a
% different theory, we may be able to re-use tactics and other forms of
% meta-programming across both.

% Existing theory exploration systems have been successfully applied to this
% problem, however the use of pure exploration misses opportunities to
% \emph{focus} the search, since we know which lemmas are used in those theories
% where a technique succeeded. If we can find an analogy to map from such solved
% problems to our unsolved goal, we can infer the approximate form of the lemmas
% we require, and target these specifically.

% The approach taken by \textsc{ACL2(ml)} is to find lemmas which may be relevant
% to solving a goal $G$ by making analogies via unsupervised clustering
% \cite{Heras.Komendantskaya.Johansson.ea:2013}. These clusters are used in two
% ways:

% \begin{itemize}
% \item First, we use the cluster $C_G$ containing $G$ to identify analogous
%   theorems.

% \item For each theorem $T \in C_G \setminus \{G\}$, we consider those symbols
%   $S_T$ which occur in $T$ but not in $G$. Our analogous lemmas are those used
%   to prove $T$, mutated such that symbols $s \in S_T$ are replaced by members of
%   the cluster $C_s$ containing $s$.
% \end{itemize}

% The running examples for demonstrating \textsc{ACL2(ml)} are equivalence
% theorems for tail-recursive and non-tail-recursive calculations, as well as the
% effect of repeating certain list operations:

% \begin{itemize}

%   \item $\forall n, \texttt{natp}(n) \rightarrow \texttt{fact-tail}(n) = \texttt{fact}(n)$ where \texttt{natp} is the predicate that $n$ is a natural number, whilst \texttt{fact-tail} and \texttt{fact} are tail-recursive and non-tail-recursive implementations of factorial, respectively.

%   \item $\forall n, \texttt{natp}(n) \rightarrow \texttt{power-tail}(n) = \texttt{power}(n)$,  where \texttt{power-tail} and \texttt{power} calculate powers of 2.

%   \item $\forall n, \texttt{natp}(n) \rightarrow \texttt{fib-tail}(n) = \texttt{fib}(n)$,  where \texttt{fib-tail} and \texttt{fib} calculate fibonacci numbers.

%   \item $\forall x, \texttt{nat-listp}(x) \rightarrow \texttt{sort}(\texttt{sort}(x)) = \texttt{sort}(x)$, for list-of-natural-numbers predicate \texttt{nat-listp} and list-sorting function \texttt{sort}.

%   \item $\forall x, \texttt{true-listp}(x) \rightarrow \texttt{rev}(\texttt{rev}(x)) = x$, where \texttt{true-listp} ensures that $x$ is a valid singly-linked list structure and \texttt{rev} is list reversal.

%   \item $\forall x, \texttt{true-listp}(x) \rightarrow \texttt{int}(x, x) = x$, where \texttt{int} is the intersection of lists (i.e. a list of elements common to each).

% \end{itemize}

% \subsection{Auxiliary Lemmas} \label{sec:auxiliarylemmas}

% One consideration when generating conjectures is the difference between
% theorems, lemmas, corollaries, etc. From a logical point of view, these are all
% equivalent, and hence most proof assistants do not distinguish between
% them. However, their \emph{intention} may be different: in a sense, theorems are
% the interesting results; whilst lemmas are useful results, required for proving
% the theorems.

% Some systems, like Coq, allow users to \emph{label} each statement as being a
% \coq{Theorem}, a \coq{Lemma}, etc. despite their internal representations being
% the same. This shows us immediately that lemmas outnumber theorems; in the Coq
% standard library there are over five times as many lemmas as theorems
% \footnote{The latest version as of writing is \texttt{coq-8.4pl6} which, when
%   excluding comments, includes 1492 occurences of \coq{Theorem} and 7594 of
%   \coq{Lemma} in its \texttt{theories/} directory.}.

% % TODO: Analyse them

% % TODO: Theory exploration as lemma generation; give example from a HipSpec paper

% We can find a need for auxiliary lemmas, once again, in the context of
% tail-recursive functions. Consider proving the (pointwise) equality of the
% following Coq functions, defined for the Peano naturals \coq{Z} and \coq{S}:

% \begin{coqblock}
% Inductive Nat : Set := Z : Nat
%                      | S : Nat -> Nat.

% Fixpoint plus      (n m : Nat) := match n with
%                                       | Z    => m
%                                       | S n' => S (plus n' m)
%                                   end.

% Fixpoint plus_tail (n m : Nat) := match n with
%                                       | Z    => m
%                                       | S n' => plus_tail n' (S m)
%                                   end.
% \end{coqblock}

% \begin{haskell}
% Haskell equivalent:

% plus :: Nat -> Nat -> Nat
% plus      n  Z    = n
% plus      n (S m) = S (plus n m)

% plus_tail :: Nat -> Nat -> Nat
% plus_tail n  Z    = n
% plus_tail n (S m) = plus_tail (S n) m
% \end{haskell}

% Both of these functions implement addition, but the \coq{plus_tail} variant is
% tail-recursive. However, if we want to \emph{prove} that the definitions are
% (pointwise) equal, we run into difficulties. In particular, when the inductive
% step requires us to prove \coq{plus (S n) m = plus n (S m)} (which seems
% reasonable), we cannot make this go through using another inductive argument.

% \begin{coqblock}
% (* Solve equalities by beta-normalising both sides *)
% Ltac triv := try (simpl; reflexivity).

% (* Prove equivalence of plus and plus_tail *)
% Theorem equiv : forall n m, plus n m = plus_tail n m.
%   induction n; triv. (* Base case is trivial *)

%   (* Inductive case: plus (S n) m = plus_tail (S n) m *)
%   intro m.

%   (* Beta-reduce the right-hand-side (justification is trivial) *)
%   replace (plus_tail (S n) m) with (plus_tail n (S m)); triv.

%   (* Use induction hypothesis to replace plus_tail with plus *)
%   rewrite <- (IHn (S m)).
% \end{coqblock}

% Specifically, the \emph{conclusion} of a second inductive hypothesis is exactly
% the equation we need:

% \begin{coqblock}
% IHn' : (forall x, plus n' x = plus_tail n' x) -> plus (S n') m = plus n' (S m)
% \end{coqblock}

% Yet we cannot provide it with the argument it needs, as our original induction
% hypothesis is \emph{too specific} (i.e. it has too many \coq{S} constructors):

% \begin{coqblock}
% IHn : forall x, plus (S n') x = plus_tail (S n') x
% \end{coqblock}

% We are forced to abandon the proof, despite such a reasonable-looking
% intermediate goal.

% In fact, if we attempt to prove that goal \emph{separately}, we can use a
% straightforward argument by induction; even though it is actually
% \emph{stronger} due to the absence of the \coq{IHn} assumption. Using this
% separate result as a lemma, the pointwise equality is proven easily.

% \begin{coqblock}
% Lemma gen n m : plus (S n) m = plus n (S m).
%   induction n; triv. (* Base case is trivial *)

%   (* Move all S constructors outside *)
%   simpl. rewrite <- IHn. simpl.

%   (* Trivial *)
%   reflexivity.
% Defined.
% \end{coqblock}

% \begin{coqblock}
%   rewrite (gen n m).
%   reflexivity.
% Defined.
% \end{coqblock}