### 2019-06-18 PhDThesis/clustering.tex
Lots of clustering changes

\chapter{Similarity and Clustering}
\label{sec:recurrentclustering}

To improve the performance of theory exploration tools, we need to avoid
brute-force search in favour of a more informed strategy. This, in turn,
requires more information about the terms in the given signature. One
generally available source of information is the syntax of a term's definition.

Unfortunately these definitions are over-complicated due to the relatively large
syntax of Haskell, which includes many convenience features (``syntactic
sugar'') and alternative ways to express the same operation (e.g. binding using
\hs{let} or \hs{where}, pattern matching with \hs{if} or \hs{case}, function
abstraction via \hs{f x = ...} or \hs{f = \x -> ...}, etc.). This obscures the
simplicity of the lambda calculus underlying the language, and more importantly
it makes syntactic analysis less useful for our purposes by drawing distinctions
between semantically identical terms. Since theory exploration is concerned only
with semantics (e.g. telling us how code actually behaves), we perform our
analyses on the syntax of the \emph{GHC Core} language instead.

\section{GHC Core}\label{sec:core}
\label{sec:ghccore}

\begin{figure}
  \begin{equation*}
    \begin{split}
      expr\    \rightarrow\ & \CVar\ id                          \\
                         |\ & \CLit\ literal                     \\
                         |\ & \CApp\ expr\ expr                  \\
                         |\ & \CLam\ \mathcal{L}\ expr           \\
                         |\ & \CLet\ bind\ expr                  \\
                         |\ & \CCase\ expr\ \mathcal{L}\ [alt]   \\
                         |\ & \CType                             \\
      id\      \rightarrow\ & \CLocal\       \mathcal{L}         \\
                         |\ & \CGlobal\      \mathcal{G}         \\
                         |\ & \CConstructor\ \mathcal{D}         \\
      literal\ \rightarrow\ & \CLitNum\ \mathcal{N}              \\
                         |\ & \CLitStr\ \mathcal{S}              \\
      alt\     \rightarrow\ & \CAlt\ altcon\ expr\ [\mathcal{L}] \\
      altcon\  \rightarrow\ & \CDataAlt\ \mathcal{D}             \\
                         |\ & \CLitAlt\ literal                  \\
                         |\ & \CDefault                          \\
      bind\    \rightarrow\ & \CNonRec\ binder                   \\
                         |\ & \CRec\ [binder]                    \\
      binder   \rightarrow\ & \CBind\ \mathcal{L}\ expr
    \end{split}
  \end{equation*}
  Where:
  \begin{tabular}[t]{l @{ $=$ } l}
    $\mathcal{S}$ & string literals    \\
    $\mathcal{N}$ & numeric literals   \\
    $\mathcal{L}$ & local identifiers  \\
    $\mathcal{G}$ & global identifiers \\
    $\mathcal{D}$ & constructor identifiers
  \end{tabular}

  \caption{Simplified syntax of GHC Core in BNF style. $[]$ and $(,)$ denote
    repetition and grouping, respectively. Other forms of literal, like machine
    words, are treated in a similar way and have been omitted for brevity.}
  \label{fig:coresyntax}
\end{figure}

GHC Core is an intermediate language of the \textsc{GHC} compiler based on
\fc{}, and described in detail in~\cite[Appendix C]{sulzmann2007system}. We
ignore type annotations and coercions, since these have no effect on the
semantics (they have no ``computational content''). The resulting sub-set of
Core\footnote{As of GHC version 7.10.2.} is shown in
Figure~\ref{fig:coresyntax}.

\begin{figure}
  \begin{subfigure}[haskell]{\textwidth}
    \underline{\texttt{Haskell definitions}}
    \begin{haskell}
plus    Z  y = y
plus (S x) y = S      (plus x y)

mult    Z  y = Z
mult (S x) y = plus y (mult x y)

odd     Z  = False
odd  (S n) = even n

even    Z  = True
even (S n) = odd n
    \end{haskell}
    \smallskip
  \end{subfigure}
  \begin{subfigure}[plus]{\textwidth}
    \begin{small}
      \underline{\texttt{plus}}
      \begin{verbatim}
Lam "a" (Lam "y" (Case (Var (Local "a"))
                       "b"
                       (Alt (DataAlt "Z") (Var (Local "y")))
                       (Alt (DataAlt "S") (App (Var (Constructor "S"))
                                               (App (App (Var (Global "plus"))
                                                         (Var (Local  "x")))
                                                    (Var (Local "y"))))
                                          "x")))
      \end{verbatim}
    \end{small}
  \end{subfigure}
  \begin{subfigure}[mult]{\textwidth}
    \begin{small}
      \underline{\texttt{mult}}
      \begin{verbatim}
Lam "a" (Lam "y" (Case (Var (Local "a"))
                       "b"
                       (Alt (DataAlt "Z") (Var (Constructor "Z")))
                       (Alt (DataAlt "S") (App (App (Var (Global "plus"))
                                                    (Var (Local  "y")))
                                               (App (App (Var (Global "mult"))
                                                         (Var (Local  "x")))
                                                    (Var (Local  "y"))))
                                          "x")))
      \end{verbatim}
    \end{small}
  \end{subfigure}
  \begin{subfigure}[odd]{\textwidth}
    \begin{small}
      \underline{\texttt{odd}}
      \begin{verbatim}
Lam "a" (Case (Var (Local "a"))
              "b"
              (Alt (DataAlt "Z") (Var (Constructor "False")))
              (Alt (DataAlt "S") (App (Var (Global "even"))
                                      (Var (Local  "n")))
                                 "n"))
      \end{verbatim}
    \end{small}
  \end{subfigure}
  \begin{subfigure}[even]{\textwidth}
    \begin{small}
      \underline{\texttt{even}}
      \begin{verbatim}
Lam "a" (Case (Var (Local "a"))
              "b"
              (Alt (DataAlt "Z") (Var (Constructor "True")))
              (Alt (DataAlt "S") (App (Var (Global "odd"))
                                      (Var (Local  "n")))
                                 "n"))
      \end{verbatim}
    \end{small}
  \end{subfigure}
  \caption{Example Haskell functions and their translations into the Core syntax
    of Figure \ref{fig:coresyntax}. Notice the introduction of explicit
    $\lambda$ abstractions (\texttt{Lam}) and the use of \texttt{Case} to
    represent piecewise definitions. Fresh variables are chosen arbitrarily
    as \texttt{"a"}, \texttt{"b"}, etc.}
  \label{fig:coreexample}
\end{figure}

GHC's translation from Haskell to Core is routine: Figure~\ref{fig:coreexample}
shows the Core representation of some simple functions. Although Core is more
verbose, it has fewer redundant features than the full Haskell syntax: for
example all pattern-matching is represented with \texttt{Case}, all function
abstraction uses \texttt{Lam}(bda), and so on. We can see that similar structure
in the Haskell definitions gives rise to similar structure in the Core, such as
the definitions of \hs{odd} and \hs{even} being identical except for the
identifiers. It is this close correspondence which allows us to analyse Core
expressions in place of their more complicated Haskell source.

As a further simplification, we also replace data constructors in definitions
with normal functions, e.g. replacing occurrences of a constructor
\hs{S :: Nat -> Nat} with a function \hs{s} defined as \hs{s x = S x} (this is
known as the $\eta$-expansion of \hs{S}). We do this to remove more distinctions
from the syntax, and also to reduce cross-language differences (in particular,
the Isabelle language used by the \isascheme{}, \isacosy{} and \hipster{} theory
exploration systems treats constructors in a different way to Haskell, so having
constructors in our data would complicate comparisons).

\section{Feature Extraction}

Raw syntax trees, even in Core form, are difficult to analyse due to their
recursive structure, use of discrete labels and heavy reliance on references.
This makes them unsuitable for many machine learning algorithms, which instead
require a fixed-size vector of continuous numbers known as a
\emph{feature vector}.

The process of turning raw data into feature vectors is called \emph{feature
  extraction}, and we adapt the methodology of \emph{recurrent clustering}
proposed in~\cite{DBLP:journals/corr/HerasK14,heras2013proof}, since this
handles recursive tree structures, resolves references and produces numeric
values. We suggest a new recurrent clustering and feature extraction algorithm
for Haskell (shown in Algorithm \ref{alg:recurrent}) and compare its similarity
and differences to the existing approaches of \textsc{ML4PG} and
\textsc{ACL2(ml)}.

We describe our algorithm in two stages: the first transforms the nested
structure of expressions into a flat vector representation; the second converts
the discrete symbols of Core syntax into features (real numbers), which we will
denote using the function $\phi$. Before introducing our algorithm, we briefly
describe \emph{clustering}, and its use in \emph{recurrent clustering}.

\subsection{Clustering}

Clustering is an unsupervised machine learning task for grouping $n$ data points
using a similarity metric. There are many variations on this theme, but in our
case we make the following choices:

\begin{itemize}
\item For simplicity, we use the ``rule of thumb'' given in
  \cite[pp. 365]{mardia1979multivariate} to fix the number of clusters at
  $k = \lceil \sqrt{\frac{n}{2}} \rceil$.
\item Data points will be $d$-dimensional feature vectors, as defined above.
\item We will use euclidean distance (denoted $e$) as our similarity metric.
\item We will use \emph{k-means} clustering, implemented by Lloyd's algorithm
  \cite{lloyd1982least}.
\end{itemize}

We choose k-means because it is a standard approach and we are more concerned
with the application of feature extraction to Haskell and its use in theory
exploration, rather than precise tuning of learning algorithms. There are many
other clustering algorithms we could use, such as \emph{expectation
  maximisation}, but experiments with \textsc{ML4PG} have shown little
difference in their results; it seems that the quality of the features is the
bottleneck to learning, so there is no reason to avoid a fast algorithm like
k-means. We use a Haskell implementation provided by the package
\hPackage{k-means}.

Since k-means is iterative, we will use function notation to denote time steps,
so $x(t)$ denotes the value of $x$ at time $t$. We denote the clusters as $C^1$
to $C^k$. As the name suggests, k-means uses the mean value of each cluster,
which we denote as $\vect{m}^1$ to $\vect{m}^k$, hence:

\begin{equation*}
  m^i_j(t) = \mean{C^i_j}(t) = \frac{\sum_{\vect{x} \in C^i(t)} x_j}{|C^i(t)|}
    \quad \text{for $t > 0$}
\end{equation*}

Before k-means starts, we must choose \emph{seed} values for
$\vect{m}^i(0)$. Many methods have been proposed for choosing these values
\cite{arthur2007k}. For simplicity, we will choose values randomly from our
data set $S$; this is known as the Forgy method.

The elements of each cluster $C^i(t)$ are those data points closest to the mean
value at the previous time step:

\begin{equation*}
  C^i(t) = \{ \vect{x} \in S \mid i = \argmin\limits_j e(\vect{x},
                                                        \vect{m}^j(t-1)) \}
    \quad \text{for $t > 0$}
\end{equation*}

As $t$ increases, the clusters $C^i$ move from their initial location around the
``seeds'', to converge on a local minimum of the ``within-cluster sum of squared
error'' objective:

\begin{equation} \label{eq:kmeansobjective}
  \argmin\limits_C \sum_{i=1}^k \sum_{\vect{x} \in C^i} e(\vect{x}, \vect{m}^i)^2
\end{equation}

Many conservative improvements can be made to the standard k-means algorithm
described above. For example, a more efficient approach like
\emph{yinyang k-means}~\cite{conf/icml/DingZSMM15} could make larger input sizes
more practical to cluster; the \emph{k-means++}
approach~\cite{arthur2007k, bahmani2012scalable} can be used to more carefully
select the ``seed'' values for the first timestep; and the \emph{x-means}
algorithm~\cite{pelleg2000x} is able to estimate how many clusters to use
(instead of relying on our ``rule of thumb''). However, like the choice of
k-means versus other clustering algorithms, these algorithmic improvements would
need higher quality data than we are capable of producing in these early
experiments.

\subsubsection{Recurrent Clustering}

\emph{Recurrent clustering} has been implemented in the
ML4PG tool for analysing Coq proofs~\cite{journals/corr/abs-1212-3618} and
\textsc{ACL2(ml)} for ACL2~\cite{heras2013proof}. Both approaches transform
syntax trees into fixed-size matrices, then concatenate the rows of these
matrices to produce feature vectors. \textsc{ACL2(ml)} produces matrices by
\emph{summarising} information about the tree; for example, one column counts
the number of variables appearing at each tree level, others count the number of
function symbols which are nullary, unary, binary, etc. In contrast, the
algorithm of ML4PG converts tokens of the syntax tree directly into numbers,
arranging them in the matrix based on their position in the syntax tree. It is
this latter algorithm that we adapt for analysing GHC Core expressions.

The key to recurrent clustering is its treatment of identifiers. From a purely
symbolic perspective, they are atomic: two identifiers can either be identical
or distinct. Consider the \hs{odd} and \hs{even} functions in
Figure~\ref{fig:coreexample}, which differ only in their choice of identifiers.
These functions are clearly different (opposite, in fact), so their identifiers
should play a role in determining their similarity. Yet these distinctions, e.g.
between \hs{True} and \hs{False}, seem intuitively small relative to the
distinction between, for example, \hs{isEmpty} and \hs{isValidEmailAddress}.

Recurrent clustering solves this dichotomy by introducing a finer granularity to
the treatment of identifiers, by first clustering the definitions that they
refer to. If two definitions appear in the same cluster, their identifiers will
be denoted by similar numbers during subsequent feature extraction; if they
appear in different clusters, their numbers will differ more during feature
extraction. It is this use of clustering to determine features that makes the
approach \emph{recurrecnt}.

\subsection{Expressions to Vectors}
\label{sec:expressionstovectors}

The first step of our feature extraction algorithm is to convert the tree
structure of Core expressions to a flat vector. Since k-means clustering
considers the elements of a feature vector to be \emph{orthogonal}, we must
ensure that similar expressions not only give rise to similar numerical values,
but crucially that these values appear \emph{at the same position} in the
feature vectors. Different patterns of nesting can alter the ``shape'' of
expressions, so simple traversals (breadth-first, depth-first, post-order, etc.)
may cause features from equivalent sub-expressions to be mis-aligned. For
example, consider the following Core expressions, which represent pattern-match
clauses with different patterns but the same body (\hs{\vlocal{y}}):

\begin{equation}\label{eq:xy}
  \begin{array}{r@{}l@{}l@{}}
    X\ &=\ \CAlt\ (\CDataAlt\ \id{C})\ & (\vlocal{y}) \\
    Y\ &=\ \CAlt\ \CDefault\           & (\vlocal{y})
  \end{array}
\end{equation}

If we traverse these expressions in breadth-first order, converting each token
to a feature using some function $\phi$ (defined later) and padding to the same
length with $0$, we would get the following feature vectors:

\begin{small}
  \begin{equation}
    \begin{array}{r@{}l@{}l@{}l@{}l@{}l@{}l@{}l}
      breadthFirst(X)\ &=\ (\feature{\CAlt},\
                           &\feature{\CDataAlt},\
                           &\feature{\CVar},\
                           &\feature{\id{C}},\
                           &\feature{\CLocal},\
                           &\feature{\id{y}} &) \\
      breadthFirst(Y)\ &=\ (\feature{\CAlt},\
                           &\feature{\CDefault},\
                           &\feature{\CVar},\
                           &\feature{\CLocal},\
                           &\feature{\id{y}},\
                           &0 &)
    \end{array}
  \end{equation}
\end{small}

Here the features corresponding to the common sub-expression $\CLocal\ \id{y}$
are misaligned, such that only $\frac{1}{3}$ of features are guaranteed to match
(others may match by coincidence, depending on $\phi$). These feature vectors
might be deemed very dissimilar during clustering, despite the intuitive
similarity of the expressions $X$ and $Y$ from which they derive.

If we were to align these features optimally, by padding the fourth column
rather than the sixth, then $\frac{2}{3}$ of features would be guaranteed to
match, making the similarity of the vectors more closely match our intuition and
depend less on coincidence.

The method we use to ``flatten'' expressions, described below, is a variation of
breadth-first traversal which pads each level of nesting to a fixed size $c$
(for \emph{columns}). This doesn't guarantee alignment, but it does prevent
mis-alignment from accumulating across different levels of nesting. Our method
would align these features into the following vectors, if $c = 2$:

\begin{small}
  \begin{equation}\label{eq:flattened}
    \begin{array}{r@{}l@{}l@{}l@{}l@{}l@{}l@{}l@{}l@{}l}
      featureVec(X)\ &=\ (\feature{\CAlt},\
                         &0,\
                         &\feature{\CDataAlt},\
                         &\feature{\CVar},\
                         &\feature{\id{C}},\
                         &\feature{\CLocal},\
                         &\feature{\id{y}},\
                         &0 &) \\
      featureVec(Y)\ &=\ (\feature{\CAlt},\
                         &0,\
                         &\feature{\CDefault},\
                         &\feature{\CVar},\
                         &\feature{\CLocal},\
                         &0,\
                         &\feature{\id{y}},\
                         &0 &)
    \end{array}
  \end{equation}
\end{small}

Here $\frac{1}{2}$ of the original 6 features align, which is more than
$breadthFirst$ but not optimal. Both vectors have also been padded by an extra 2
zeros compared to $breadthFirst$; raising their alignment to $\frac{5}{8}$.

\begin{figure}
  \centering
  \begin{scriptsize}
      \Tree[ .$\feature{\CLam}$
                $\feature{\id{a}}$
                [ .$\feature{\CCase}$
                     [ .$\feature{\CVar}$
                          [ .$\feature{\CLocal}$
                               $\feature{\id{a}}$ ]]
                     $\feature{\id{b}}$
                     [ .$\feature{\CAlt}$
                          $\feature{\CDataAlt}$
                          [ .$\feature{\CVar}$
                               $\feature{\CConstructor}$ ]]
                     [ .$\feature{\CAlt}$
                          $\feature{\CDataAlt}$
                          [ .$\feature{\CApp}$
                               [ .$\feature{\CVar}$
                                    [ .$\feature{\CGlobal}$
                                         $\feature{\id{even}}$ ]]
                               [ .$\feature{\CVar}$
                                    [ .$\feature{\CLocal}$
                                         $\feature{\id{n}}$ ]]]
                          $\feature{\id{n}}$ ]]]
  \end{scriptsize}
  \caption[Rose tree for odd]{\label{fig:rosetreeexample} Rose tree for the
    expression \hs{odd} from Figure \ref{fig:coreexample}. Each (sub-) rose tree
    is rendered with its feature at the node and sub-trees beneath.}
\end{figure}

To perform this flattening we first transform the nested tokens of an expression
into a \emph{rose tree} of features, via the function $toTree$ defined below.
Intuitively, rose trees are simply an s-expression representation of the Core
syntax, with the feature extraction function $\phi$ mapped over the labels:
Figure~\ref{fig:rosetreeexample} shows the result of transforming the \hs{odd}
function from Figure~\ref{fig:coreexample}.

We follow the presentation in~\cite{blundell2012bayesian} and define rose trees
recursively as follows: $T$ is a rose tree if $T = (f, T_1, \dots, T_{n_T})$,
where $f \in \mathbb{R}$ and $T_i$ are rose trees.

$T_i$ are the \emph{sub-trees} of $T$ and $f$ is the \emph{feature at} $T$.

$n_T$ may differ for each (sub-) tree; trees where $n_T = 0$ are \emph{leaves}.

The recursive definition of $toTree$ is mostly routine; each repeated element
(shown as $\dots$) has an example to indicate their handling, e.g. for $\CRec$
we apply $toTree$ to each $e_i$. We ignore values of $\mathcal{D}$, since
constructors have no internal structure for us to compare. We also ignore values
from $\mathcal{S}$ and $\mathcal{N}$ as it simplifies our later definition of
$\phi$, and we conjecture that the effect of such ``magic values'' on clustering
real code is low.

\begin{align*}\label{fig:totree}
  toTree(e) &=
  \begin{cases}
    (\feature{\CVar},     toTree(e_1))                                 & \text{if $e = \CVar\ e_1$} \\
    (\feature{\CLit},     toTree(e_1))                                 & \text{if $e = \CLit\ e_1$} \\
    (\feature{\CApp},     toTree(e_1), toTree(e_2))                    & \text{if $e = \CApp\ e_1\ e_2$} \\
    (\feature{\CLam},     toTree(e_1))                                 & \text{if $e = \CLam\ l_1\ e_1$} \\
    (\feature{\CLet},     toTree(e_1), toTree(e_2))                    & \text{if $e = \CLet\ e_1\ e_2$} \\
    (\feature{\CCase},    toTree(e_1), toTree(a_1), \dots)             & \text{if $e = \CCase\ e_1\ l_1\ a_1\ \dots$} \\
    (\feature{\CType})                                                & \text{if $e = \CType$} \\
    (\feature{\CLocal},   (\feature{l_1}))                            & \text{if $e = \CLocal\ l_1$} \\
    (\feature{\CGlobal},  (\feature{g_1}))                            & \text{if $e = \CGlobal\ g_1$} \\
    (\feature{\CConstructor})                                         & \text{if $e = \CConstructor\ d_1$} \\
    (\feature{\CLitNum})                                              & \text{if $e = \CLitNum\ n_1$} \\
    (\feature{\CLitStr})                                              & \text{if $e = \CLitStr\ s_1$} \\
    (\feature{\CAlt},     toTree(e_1), toTree(e_2))                   & \text{if $e = \CAlt\ e_1\ e_2\ l_1\ \dots$}  \\
    (\feature{\CDataAlt})                                             & \text{if $e = \CDataAlt\ g_1$}  \\
    (\feature{\CLitAlt},  toTree(e_1))                                & \text{if $e = \CLitAlt\ e_1$}  \\
    (\feature{\CDefault})                                             & \text{if $e = \CDefault$}  \\
    (\feature{\CNonRec},  toTree(e_1))                                & \text{if $e = \CNonRec\ e_1$}  \\
    (\feature{\CRec},     toTree(e_1), \dots)                         & \text{if $e = \CRec\ e_1\ \dots$} \\
    (\feature{\CBind},    toTree(e_1))                                & \text{if $e = \CBind\ l_1\ e_1$}
  \end{cases}
\end{align*}

\begin{figure}
    \begin{equation*}
      \begin{bmatrix}
        \feature{\CLam}      & 0                       & 0                 & 0                   & 0               & 0                \\
        \feature{\id{a}}     & \feature{\CCase}        & 0                 & 0                   & 0               & 0                \\
        \feature{\CVar}      & \feature{\id{b}}        & \feature{\CAlt}   & \feature{\CAlt}     & 0               & 0                \\
        \feature{\CLocal}    & \feature{\CDataAlt}     & \feature{\CVar}   & \feature{\CDataAlt} & \feature{\CApp} & \feature{\id{n}} \\
        \feature{\id{a}}     & \feature{\CConstructor} & \feature{\CVar}   & \feature{\CVar}     & 0               & 0                \\
        \feature{\CGlobal}   & \feature{\CLocal}       & 0                 & 0                   & 0               & 0                \\
        \feature{\id{even}}  & \feature{\id{n}}        & 0                 & 0                   & 0               & 0
      \end{bmatrix}
    \end{equation*}
    \caption{Matrix generated from Figure \ref{fig:rosetreeexample}, padded to 6 columns. Each level of nesting in the tree corresponds to a row in the matrix.}
    \label{fig:matrixexample}
\end{figure}

These rose trees are then turned into matrices, as shown in
Figure~\ref{fig:matrixexample}. Each row $i$ of the matrix contains the features
at depth $i$ in the rose tree, read left-to-right, followed by any required
padding. These matrices are then either truncated, or padded with rows (on the
bottom) or columns (on the right) of zeros, to fit a fixed size $r \times c$.

\begin{sloppypar}
  Finally, matrices are turned into vectors by concatenating the rows from top
  to bottom, hence Figure~\ref{fig:matrixexample} will produce a vector
  beginning
  $(\feature{\CLam} ,
    0               ,
    0               ,
    0               ,
    0               ,
    0               ,
    \feature{\id{a}},
    \feature{\CCase},
    0               ,
    0               ,
    0               ,
    0               ,
    \feature{\CVar} ,
    \feature{\id{b}},
    \feature{\CAlt} ,
    \feature{\CAlt} ,
    \dots$.
\end{sloppypar}

\subsubsection{Symbols to Features}
\label{sec:symbolstofeatures}

We now define the function $\phi$, which turns terminal symbols of Core syntax
into features (real numbers). For known language features, such as
$\feature{\CLam}$ and $\feature{\CCase}$, we can enumerate the possibilities and
assign a value to each, in a similar way to~\cite{DBLP:journals/corr/HerasK14}
in Coq. We use a constant $\alpha$ to separate these values from those of other
tokens (e.g. identifiers), but the order is essentially
arbitrary:~\footnote{In~\cite{DBLP:journals/corr/HerasK14}, ``similar'' Gallina
  tokens like \coq{fix} and \coq{cofix} are grouped together to reduce
  redundancy; we do not group tokens, but we do put ``similar'' tokens close
  together, such as \CLocal\ and \CGlobal.}

\begin{equation} \label{eq:feature}
  \begin{aligned}
    \feature{\CAlt}          &= \alpha      &
    \feature{\CDataAlt}      &= \alpha + 1  &
    \feature{\CLitAlt}       &= \alpha + 2  \\
    \feature{\CDefault}      &= \alpha + 3  &
    \feature{\CNonRec}       &= \alpha + 4  &
    \feature{\CRec}          &= \alpha + 5  \\
    \feature{\CBind}         &= \alpha + 6  &
    \feature{\CLet}          &= \alpha + 7  &
    \feature{\CCase}         &= \alpha + 8  \\
    \feature{\CLocal}        &= \alpha + 9  &
    \feature{\CGlobal}       &= \alpha + 10 &
    \feature{\CConstructor}  &= \alpha + 11 \\
    \feature{\CVar}          &= \alpha + 12 &
    \feature{\CLam}          &= \alpha + 13 &
    \feature{\CApp}          &= \alpha + 14 \\
    \feature{\CType}         &= \alpha + 15 &
    \feature{\CLit}          &= \alpha + 16 &
    \feature{\CLitNum}       &= \alpha + 17 \\
    \feature{\CLitStr}       &= \alpha + 18
  \end{aligned}
\end{equation}

To encode \emph{local} identifiers $\mathcal{L}$ we would like a quantity which
gives equal values for $\alpha$-equivalent expressions (i.e. renaming an
identifier shouldn't affect the feature). To do this, we use the \emph{de Bruijn
  index} of the identifier~\cite{de1972lambda}, denoted $i_l$:

\begin{equation} \label{eq:localfeature}
  \feature{l} = i_l + 2 \alpha \quad \text{if $l \in \mathcal{L}$}
\end{equation}

We again use $\alpha$ to separate these features from those of other constructs.

We discard the contents of literals and constructor identifiers when converting
to rose trees, so the only remaining case is global identifiers
$\mathcal{G}$. Since these are declared \emph{outside} the body of an
expression, we cannot perform the same indexing trick as for local
identifiers. We also cannot directly encode the form of the identifiers,
e.g. using a scheme like G{\"o}del numbering, since this is essentially
arbitrary and has no effect on their semantic meaning (referencing other
expressions).

Instead, we use the approach taken nu ML4PG and encode global identifiers
\emph{indirectly}, by looking up the expressions which they \emph{reference}:

\begin{equation} \label{eq:globalfeature}
  \feature{g \in \mathcal{G}} =
    \begin{cases}
      i + 3 \alpha \quad & \text{if $g \in C_i$} \\
      f_{recursion}         & \text{otherwise}
    \end{cases}
\end{equation}

Where $C_i$ are our clusters (in some arbitrary order). This is where the
recurrent nature of the algorithm appears: to determine the contents of $C_i$,
we must perform k-means clustering \emph{during} feature extraction; yet that
clustering step, in turn, requires that we perform feature extraction.

For this recursive process to be well-founded, we perform a topological sort on
declarations based on their dependencies (the expressions they reference). In
this way, we can avoid looking up expressions which haven't been clustered
yet. To handle mutual recursion we use Tarjan's algorithm~\cite{tarjan1972depth}
to produce a sorted list of \emph{strongly connected components} (SCCs), where
each SCC is a mutually-recursive sub-set of the declarations. If an identifier
cannot be found amongst those clustered so far, it must appear in the same SCC
as the expression we are processing; hence we give that identifier the constant
feature value $f_{recursion}$.

\begin{algorithm}
  \begin{algorithmic}[1]
    \Require List $d$ contains SCCs of (identifier, expression) pairs, in
    dependency order.
    \Procedure{RecurrentCluster}{}
      \State $\vect{C} \gets []$
      \State $DB \gets \varnothing$
      \ForAll{$scc$ \textbf{in} $d$}
        \State $DB \gets DB \cup \{(i, featureVec(e)) \mid (i, e) \in scc\}$
        \State $\vect{C} \gets kMeans(DB)$
      \EndFor
      \Return $\vect{C}$
    \EndProcedure
  \end{algorithmic}
  \caption{Recurrent clustering of Core expressions.}
  \label{alg:recurrent}
\end{algorithm}

By working through the sorted list of SCCs, storing the features of each
top-level expression as they are calculated, our algorithm can be computed
\emph{iteratively} rather than recursively, as shown in
Algorithm~\ref{alg:recurrent}.

\begin{figure}
  \centering
  \begin{subfigure}{\textwidth}
    \centering
    \begin{tikzpicture}[node distance=2cm]
      \tikzset{vertex/.style = {minimum size=1.5em}}
      \tikzset{edge/.style = {->,> = latex'}}

      \node[vertex]                at (0.5\textwidth, 3)   (z)     {\hs{Z}};
      \node[vertex]                at (0.5\textwidth, 0)   (s)     {\hs{S}};
      \node[vertex]                at (0,             1.5) (plus)  {\hs{plus}};
      \node[vertex, anchor=east]   at (\textwidth,    1.5) (false) {\hs{False}};
      \node[vertex, right of=plus]                         (mult)  {\hs{mult}};
      \node[vertex, left of=false]                         (odd)   {\hs{odd}};
      \node[vertex, left of=odd]                           (even)  {\hs{even}};
      \node[vertex, left of=even]                          (true)  {\hs{True}};

      \draw[edge, dashed]     (z)     to (plus);
      \draw[edge, dashed]     (z)     to (mult);
      \draw[edge, dashed]     (z)     to (odd);
      \draw[edge, dashed]     (z)     to (even);
      \draw[edge, dashed]     (s)     to (plus);
      \draw[edge, dashed]     (s)     to (mult);
      \draw[edge, dashed]     (s)     to (odd);
      \draw[edge, dashed]     (s)     to (even);
      \draw[edge, dashed]     (false) to (odd);
      \draw[edge, dashed]     (true)  to (even);
      \draw[edge]             (plus)  to (mult);
      \draw[edge]             (even)  to (odd);
      \draw[edge]             (odd)   to (even);
      \draw[edge, loop above] (plus)  to (plus);
      \draw[edge, loop above] (mult)  to (mult);
    \end{tikzpicture}

    \caption{Dependency graph for Figure \ref{fig:coreexample}. Loops indicate
      recursive functions, double arrows indicate mutual recursion. Dashed lines
      show references to data constructors, which we do not consider in this
      work.}
    \label{fig:dagexample}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
    \centering
    \tikzstyle{block} = [rectangle, draw]
    \tikzstyle{line}  = [draw, -latex']
    \begin{tikzpicture}[node distance=2cm]
      \node [block] (plusset) {
        \begin{tikzpicture}[node distance = 2cm]
          \node [block] (plus) {\hs{plus}};
        \end{tikzpicture}
      };
      \node [block, right of=plusset] (parityset) {
        \begin{tikzpicture}[node distance = 1cm]
          \node [block]               (odd)  {\hs{odd}};
          \node [block, below of=odd] (even) {\hs{even}};
        \end{tikzpicture}
      };
      \node [block, right of=parityset] (multset) {
        \begin{tikzpicture}[node distance = 2cm]
          \node [block] (mult) {\hs{mult}};
        \end{tikzpicture}
      };

      \path [line] (plusset)   -- (parityset);
      \path [line] (parityset) -- (multset);
    \end{tikzpicture}
    \caption{One possible topological sorting of simply connected components for
      Figure \ref{fig:dagexample} (ignoring constructors). \hs{even} and
      \hs{odd} are mutually recursive, neither can appear before the other, so
      they are grouped into one component and handled concurrently by Algorithm
      \ref{alg:recurrent}.}
    \label{fig:sccexample}
  \end{subfigure}
  \caption{Sorting functions from Figure \ref{fig:coreexample} into dependency order.}
  \label{fig:dependencyexample}
\end{figure}

As an example of this recurrent process, we can consider the Peano arithmetic
functions from Figure~\ref{fig:coreexample}. A valid topological ordering is
given in Figure~\ref{fig:sccexample}, which can be our value for $d$ (eliding
Core expressions to save space):

\begin{equation}
  d = [\{(\hs{plus}, \dots)\},
       \{(\hs{odd}, \dots), (\hs{even}, \dots)\},
       \{(\hs{mult}, \dots)\}]
\end{equation}

We can then trace the execution of Algorithm~\ref{alg:recurrent} as follows:

\begin{itemize}
\item The first iteration through \textsc{RecurrentCluster}'s loop will set
  $scc \gets \{(\hs{plus}, \dots)\}$.
\item With $i = \hs{plus}$ and $e$ as its Core expression, calculating
  $featureVec(e)$ is straightforward; the recursive call $\feature{\hs{plus}}$
  will become $f_{recursion}$ (since \hs{plus} doesn't appear in $\vect{C}$).
\item The call to $kMeans$ will produce $\vect{C} \gets [\{\hs{plus}\}]$, i.e. a
  single cluster containing \hs{plus}.
\item The next iteration will set
  $scc \gets \{(\hs{odd}, \dots), (\hs{even}, \dots)\}$.
\item With $i = \hs{odd}$ and $e$ as its Core expression, the call to \hs{even}
  will result in $f_{recursion}$.
\item Likewise for the call to $\hs{even}$ when $i = \hs{odd}$.
\item Since the feature vectors for \hs{odd} and \hs{even} will be identical,
  $kMeans$ will put them in the same cluster. To avoid the degenerate case of a
  single cluster, for this example we will assume that $k = 2$; in which case
  the other cluster must contain \hs{plus}. Their order is arbitrary, so one
  possibility is $\vect{C} = [\{\hs{odd}, \hs{even}\}, \{\hs{plus}\}]$.
\item Finally \hs{mult} will be clustered. The recursive call will become
  $f_{recursion}$ whilst the call to \hs{plus} will become $2 + 3 \alpha$, since
  $\hs{plus} \in C_2$.
\item \begin{sloppypar}Again assuming that $k = 2$, the resulting clusters will
    be
    \mbox{$\vect{C} \gets [\{\hs{odd}, \hs{even}\}, \{\hs{plus},
      \hs{mult}\}]$}.\end{sloppypar}
\end{itemize}

Even in this very simple example we can see a few features of our algorithm
emerge. For example, \hs{odd} and \hs{even} will always appear in the same
cluster, since they only differ in their choice of constructor names (which are
discarded by $toTree$) and recursive calls (which are replaced by
$f_{recursion}$).

\section{Implementation}
\label{sec:implementation}

We provide an implementation of our recurrent clustering algorithm in a tool
called \mlforhs{}. We obtain Core ASTs from Haskell packages using a plugin
for the GHC compiler~\footnote{Available at
  \url{https://github.com/warbo/ast-plugin}}, which emits a serialised version
of each Core definition as it is being compiled. This approach is more robust
than, for example, parsing source files, since it avoids the complications of
preprocessors and language extensions altering the syntax.

A post-processing stage determines which Core definitions can be used by
\quickspec{}, based on their type, visibility (whether they are encapsulated inside
their module or visible to importers), etc. For each definition, we simply use
the GHCi interpreter to check if calling \quickspec{} with that argument is
well-typed. This information, along with type signatures, arity, etc. are stored
alongside the Core definitions in JSON format.

The definitions are then sorted topologically, based on the non-local
identifiers appearing in their ASTs, and feature vectors are constructed using a
Haskell implementation of the approach described in \S
\ref{sec:recurrentclustering}. Since the features associated with each
identifier may vary between iterations (as the clusters change), we leave the
raw identifiers in the vector so their features can be extracted in the correct
context.

We implement Algorithm \ref{alg:recurrent} as a simple shell script, which
invokes Weka for clustering and associates each Core definition with a cluster
number. These numbers are used to finish the deferred feature extraction of
identifiers, the resulting feature vectors are clustered, and the process is
repeated until all SCCs have been processed.

Once the recurrent clustering is complete, we generate a string of Haskell code
for each of the final clusters, which will invoke \quickspec{} with a suitable
signature. This involves:

\begin{itemize}
\item{Monomorphising}: If a value has polymorphic type, e.g. \hs{equal :: forall
    t. t -> t -> Bool}, we must choose a concrete representation to use (in this
  case for \hs{t}), in order to know which random generator to use. We take the
  approach used in \quickcheck{} by attempting to instantiate all type variables to
  \hs{Integer}. Any values where this is invalid, such as those with
  incompatible class constraints (e.g. \hs{forall t. IsString t => t}, where
  \hs{Integer} does not implement \hs{IsString}) will not be included in the
  signature (this is checked at the AST post-processing stage).

\item{Qualifying}: All names are \emph{qualified} (prefixed by their module's
  name), to avoid most ambiguity. There is still the possibility that multiple
  packages will declare modules of the same name, although this is rare as it
  causes problems for any Haskell programmer trying to use those modules. In
  such cases the exploration process simply aborts.

\item{Appending variables}: Once a \quickspec{} theory has been defined containing
  all of the given terms, we inspect the types it references and append three
  variables for each to the theory (enough to discover laws such as
  associativity, but not too many to overflow the limit of \quickspec{}'s exhaustive
  search).

\item{Sandboxing}: One difficulty with Haskell's packaging infrastructure is
  that all required packages and modules must be provided up-front, usually by
  specification in a Cabal file. Since \mlspec{} builds signatures
  \emph{dynamically}, depending on the cluster information it is given, we do
  not know what packages it may need. To work around this problem, we evaluate
  these strings of Haskell using a custom library called \texttt{nix-eval}. This
  uses the Nix package manager to obtain all of the required packages and make
  them available to GHC, and is described further in
  Appendix~\ref{sec:nix-eval}.

\end{itemize}

The equations resulting from evaluating these strings are collected and
outputted as JSON values, to ease further processing (e.g. displaying in some
form, sending to a theorem prover, etc.).

\section{Evaluation}
\label{sec:evaluation}

% TODO: Move recurrent clustering to its own chapter, with this as its
% evaluation?
We have applied our recurrent clustering algorithm to several scenarios, with
mixed results. A major difficulty in evaluating these clusters is that we have
no ``ground truth'', i.e. there is no objectively correct way to compare
expressions. Instead, we provide a qualitative overview of the more interesting
characteristics.

As a simple example, we clustered (Haskell equivalents of) the running examples
used to present ACL2(ml) \cite{heras2013proof}, shown in Figure
\ref{fig:lisp}. These include tail-recursive and non-tail-recursive
implementations of several functions. We expect those with similar
\emph{structure} to be clustered together, rather than those which implement the
same function. The results are shown in Figure \ref{fig:haskellcluster}, where
we can see the ``\hs{Tail}'' functions clearly distinguished, with little
distinction between the tail recursive and na\"{\i}ve implementations.

Next we tested whether these same functions would be clustered together when
mixed with seemingly-unrelated functions, in this case 207 functions from
Haskell's \hs{text} library. In fact, the \hs{helperFib} and \hs{fibTail}
functions appeared together in a separate cluster from the rest. This was
unexpected, with no obvious semantic connection between these two functions and
the others in their cluster (although most are recursive, due to the nature of
the \hs{text} library).

\begin{figure}
  \begin{common-lisp}
(defun fact (n)
  (if (zp n) 1 (* n (fact (- n 1)))))

(defun helper-fact (n a)
  (if (zp n) a (helper-fact (- n 1) (* a n))))

(defun fact-tail (n)
  (helper-fact n 1))

(defun power (n)
  (if (zp n) 1 (* 2 (power (- n 1)))))

(defun helper-power (n a)
  (if (zp n) a (helper-power (- n 1) (+ a a))))

(defun power-tail (n)
  (helper-power n 1))

(defun fib (n)
  (if (zp n)
      0
      (if (equal n 1)
          1
          (+ (fib (- n 1)) (fib (- n 2))))))

(defun helper-fib (n j k)
  (if (zp n)
      j
      (if (equal n 1)
          k
          (helper-fib (- n 1) k (+ j k)))))

(defun fib-tail (n)
  (helper-fib n 0 1))
  \end{common-lisp}
  \caption{Common Lisp functions, both tail-recursive and non-tail-recursive.}
  \label{fig:lisp}
\end{figure}

\begin{figure}
  \begin{haskell}
fact n = if n == 0
            then 1
            else n * fact (n - 1)

helperFact n a = if n == 0
                    then a
                    else helperFact (n - 1) (a * n)

factTail n = helperFact n 1

power n = if n == 0
             then 1
             else 2 * power (n - 1)

helperPower n a = if n == 0
                     then a
                     else helperPower (n - 1) (a + a)

powerTail n = helperPower n 1

fib n = if n == 0
           then 0
           else if n == 1
                   then 1
                   else fib (n - 1) + fib (n - 2)

helperFib n j k = if n == 0
                     then j
                     else if n == 1
                          then k
                          else helperFib (n - 1) k (j + k)

fibTail n = helperFib n 0 1
  \end{haskell}
  \caption{Haskell equivalents of the Common Lisp functions in Figure \ref{fig:lisp}.}
  \label{fig:haskelllisp}
\end{figure}

\begin{figure}
  \centering
    \tikzstyle{block} = [rectangle, draw]
    \tikzstyle{line}  = [draw, -latex']
    \begin{tikzpicture}[node distance=3cm]
      \node [block] (c1) {
        \begin{tikzpicture}[node distance = 1cm]
          \node [block]                    (factTail)  {\hs{factTail}};
          \node [block, below of=factTail] (fibTail)   {\hs{fibTail}};
          \node [block, below of=fibTail]  (powerTail) {\hs{powerTail}};
        \end{tikzpicture}
      };
      \node [block, right of=c1] (c2) {
        \begin{tikzpicture}[node distance = 1cm]
          \node [block]                (fact)       {\hs{fact}};
          \node [block, below of=fact] (helperFact) {\hs{helperFact}};
        \end{tikzpicture}
      };
      \node [block, right of=c2] (c3) {
        \begin{tikzpicture}[node distance = 1cm]
          \node [block]                       (fib)         {\hs{fib}};
          \node [block, below of=fib]         (helperFib)   {\hs{helperFib}};
          \node [block, below of=helperFib]   (helperPower) {\hs{helperPower}};
          \node [block, below of=helperPower] (power)       {\hs{power}};
        \end{tikzpicture}
      };
  \end{tikzpicture}
  \caption{Typical clusters for the functions in Figure \ref{fig:haskelllisp}.}
  \label{fig:haskellcluster}
\end{figure}

We have also applied our recurrent clustering algorithm to a variety of the
most-downloaded packages from \hackage{} (as of 2015-10-30), including
\texttt{text} (as above), \texttt{pandoc}, \texttt{attoparsec},
\texttt{scientific}, \texttt{yesod-core} and \texttt{blaze-html}. Whilst we
expected functions with a similar purpose to appear together, such as the
various reader and writer functions of \hs{pandoc}, there were always a few
exceptions which became separate for reasons which are still unclear.

When clustering the \hs{yesod} Web framework, the clustering did seem to match
our intuitions, in particular since all 15 of Yesod's MIME type identifiers
appeared in the same cluster.

Whilst recurrent clustering has produced results which merit further
investigation, the application to theory exploration has yet to be tested
empirically. This is due to \quickspec{}'s use of \quickcheck{}'s \hs{Arbitrary} type
class to generate random values for instantiating variables. Whilst we can
automatically define \quickspec{} theories and invoke them with \hs{nix-eval}, not
all types have \hs{Arbitrary} instances; those without cannot be given any
variables in our signature, which severely limits the possible combinations
which can be explored. In many cases, no variables can be included at all,
leaving just equations involving constants. This has so far prevented us from
measuring the direct impact on \quickspec{} performance, either directly by
exploring the sub-sets identified through recurrent clustering, or indirectly by
comparing the equations generated by a full brute-force search to our recurrent
clusters: those equations relating terms from different clusters would not be
discovered by our method. It is this ratio of equations found through brute
force to those found after narrowing-down by clusters which is one of our key
objectives to maximise at this stage; until we begin to pursue the
\emph{interestingness} of the properties.

The following less-serious problems were also encountered while applying
\mlforhs{} to \hackage{} packages:

\begin{itemize}
\item Some packages, such as \texttt{warp} and \texttt{conduits}, get no
  declarations to cluster. This is because they make all of their declarations
  privately, e.g. in ``internal'' modules, then use separate modules to export
  the public declarations. GHC's renaming phase makes all references to such
  exports canonical, by pointing them to the private declarations. This forces
  us to ignore such declarations, as \quickspec{} will not be able to access them.

\item Since we do not support type-level entities, we ignore type
  classes. Unfortunately, this also means ignoring any value-level bindings (AKA
  ``methods'') which occur in a type class instance. Instead of being clustered,
  these result in references getting $f_{recursion}$ features. This is
  especially noticable in libraries like \hs{scientific}, where only the
  functions for constructing and destructing numbers in scientific notation are
  clustered; all of the arithmetic is defined in type classes. One difficulty
  with supporting methods is that their namespace in Core is disjoint from that
  of regular Haskell identifiers: a transformation layer would be required,
  along with explicit type annotations to avoid ambiguity.
\end{itemize}

It seems like this recurrent clustering method has promise, although it will
require a more thorough exploration of the parameters to obtain more intuitively
reliable results. These clusters can then be used in several ways to perform
theory exploration; the most na\"{\i}ve way being to explore each cluster as
\quickspec{} signature in its own right. \mlforhs{} already provides this
functionality, although the lack of test generators severely limits what can be
discovered.

As alluded to previously, we also have the opportunity to reason by
\emph{analogy}. Similar to the work on ACL2(ml) \cite{heras2013proof}, we could
produce a general ``scheme'' from each equation we find (either through \quickspec{}
or by data mining test suites); like Isabelle approaches have shown
\cite{Montano-Rivas.McCasland.Dixon.ea:2012}, these schemes could then be
instantiated to a variety of similar values, in an attempt to find new theorems
which are analogues of existing results from a different context. ``Mutating''
existing theorem statements in such a way would also increase the chance of any
result being considered interesting; since it's likely that the unmodified
statement was deemed interesting, and the new result would not in general follow
as a simple logical consequence.

\subsection{Recurrent Clustering}
\label{sec:clusteringexpressions}

Our recurrent clustering algorithm is similar to that of ML4PG, as our
transformation maps the elements of a syntax tree to distinct cells in a
matrix. In contrast, the matrices produced by ACL2(ml) \emph{summarise} the tree
elements: providing, for each level of the tree, the number of variables,
nullary symbols, unary symbols, etc.

Unlike ML4PG, our algorithm can handle mutually-recursive definitions, but it
also ignores types. This is because obtaining the type of each component of a
Core definition is more difficult than in ML4PG (which queries the current
\textsc{Proof General} state), and is left for future work.

The way we \emph{use} our clusters to inform theory exploration is actually more
similar to that of ACL2(ml) than ML4PG. ML4PG can either present clusters to the
user for inspection, or produce automata for recreating proofs. In ACL2(ml), the
clusters are used to restrict the search space of a proof search, much like we
restrict the scope of theory exploration.

\section{Conclusion and Future Work}
\label{sec:conclusion}

Our use of clustering to pre-process \quickspec{} signatures has required many
decisions and tradeoffs to be made. Hence our approach is just one possibility
out of many alternatives which could be investigated to push this work further.

The most obvious next step is to incorporate types. Types contain valuable
information about an expression, and would allow us to distinguish between
constructors. Since our algorithm closely follows that of ML4PG, which
\emph{does} support types, the only barrier is the practical issue of
propagating annotations through all of the Core definitions.

More speculative directions include the use of \emph{learned} representations,
rather than our hand-crafted features \cite{bengio2013representation}. This
would provide an interesting comparison, as well as being more robust in the
face of language evolution. Another intriguing possibility is to extend the
recurrent nature of our algorithm to make use of the discovered properties
during clustering; for example, by treating the discovered equations as rewrite
rules to reduce the ASTs prior to feature extraction.