### 2019-10-16 PhDThesis/litreview.tex
TODOs

\chapter{Literature Review}

\iffalse TODO: Alison: Each chapter should have a description at the start (see
bucketing ``in this chapter'') \fi

\input{litreview_automatedreasoning.tex}
\input{litreview_staticanalysis.tex}
\input{litreview_testing.tex}
\input{litreview_premiseselection.tex}
\input{litreview_machinelearning.tex}
\input{litreview_interestingness.tex}