### 2019-10-18 PhDThesis/litreview_machinelearning.tex
More literview

\section{Machine Learning Applications in Mathematics, Logic and Programming
  Language Theory}

Machine Learning (ML) applies statistical techniques to find patterns in data.
ML algorithms are numerical in nature, and often rely on assumptions such as
continuity and differentiability. In contrast, the worlds of mathematical
formulae, formal logic and computer programs are usually discrete and symbolic.
This makes application of ML to these domains more difficult than, say, image
processing.

One application of ML in these domains is to analyse real-world measurements
parameterised by some discrete entity, such as time or power used to execute
computer programs. Since flipping a single bit can cause arbitrarily large
changes in program behaviour (if it is the scrutinee of a conditional, for
example), such programs must be very similar for statistical patterns to emerge
in their executions. In the case of \emph{auto-tuning}, many programs variants
are measured, which all perform the same task in slightly different ways (e.g.
using different compiler optimisations, or dividing work differently between
hardware components); in this case the goal is to learn a predictive model of
which variant performs best for a given input~\cite{ganapathi2009case}.

Whilst performance measurements are simple numeric values, the symbolic values
need some conversion to become amenable to learning. This is usually performed
by a \emph{feature extraction} step, which takes various measurements (e.g.
counting occurrences of some element, assigning indices to different constants,
etc.) and packs them into a vector for subsequent analysis. More recently, so
called \emph{deep learning} systems have out performed those with manually
chosen features. Deep learning doesn't actually avoid feature extraction, but
its use of many ``layers'' of processing between the input and output makes the
selection of those features less important. For example, the feature extraction
in~\cite{cummins2017synthesizing} represents computer programs (after some
normalisation to a consistent layout and naming scheme) as a sequence of
individual characters, with each ASCII character replaced by a different number.
Yet even this crude method is enough for the deep learning system that consumed
these sequences to infer structure relevant for deciding whether it was better
suited to running on a CPU or GPU. The \textsc{DeepTune} system encodes slightly
more domain knowledge by tokenising language keywords, but is otherwise
similar~\cite{cummins2017end}.

\emph{Generative models} are another useful application of ML techniques to
symbolic systems. The \textsc{DeepSmith} system analyses a corpus of source code
to build a model of the programming language, which can then generate realistic
fragments of code, suitable for fuzz-testing compilers for that
language~\cite{cummins2017deepsmith}.