### 2019-11-29 PhDThesis/background.tex
A bunch of cleanup

\chapter{Background}

% Background should cover approaches to automated reasoning about software, and
% formal systems in general. We can highlight how the language/representation
% imposes invariants and guarantees on what we can know, and how we can extend
% those guarantees with features like purity (prevents behaviour depending on
% ambient environment/state), lazy evaluation (prevents problems with
% non-termination), sound type systems (limits the space of possible values for
% each expression), etc. and therefore how Haskell is a natural vehicle for us to
% use. Giving background on Haskell naturally leads to descendants like Agda and
% Idris, which are also interactive theorem provers; which lets us discuss
% Isabelle, Coq, HOL, etc., how they're difficult for existing reasoning
% algorithms to handle (e.g. resolution, which works well in first-order logics),
% approaches to helping

This chapter provides background on the fields we investigate, and the
techniques we apply. It includes descriptions of the proof- and
programming-languages we use, and justifications for their use in our work. A
very short summary of relevant Machine Learning topics is also given, to provide
context for the techniques we use (namely unsupervised clustering) and the
potential alternatives and future directions that are most applicable.

\input{background_haskell.tex}
\input{background_quickcheck.tex}
\input{background_theoryexploration.tex}

% FIXME: This seems to be repetitive of Theory Exploration?
% \input{background_conjecturegeneration.tex}

\input{background_machinelearning.tex}